---
title: "Aurelia Notebook | AAUR2 mapping"
output: html_notebook
---

#Aurelia Cell Atlas | Complete

```{r Setup, echo=F, warning=FALSE}


#load libraries: 
library(Seurat,quietly=T)
library(RColorBrewer,quietly=T)
library(patchwork,quietly=T)
library(ggplot2,quietly=T)
library(pals,quietly=T)
library(readxl,quietly=T)
library(SeuratWrappers,quietly=T)
library(tidyr,quietly=T)

```

```{r colours, echo=F}

load (file="Aaur2.newnames.RData")
genes=Aa_genes
  AaTF_list <- which(!(Aa_genes$TF_Fam=='-'))
  AaTF_list = Aa_genes$gene_short_name[AaTF_list]
LibCP =brewer.paired(10)
# pal.bands(brewer.paired(10))
lib.names = c('Aa.polyp.1','Aa.polyp.2','Aa.polyp.clover','Aa.strobila','Aa.ephyra','Aa.medusa','Aa.gastrodermis',
              'Aa.manubrium','Aa.margin','Aa.umbrella') #names were too long...
gene.cp.dark=c('grey30',rev(brewer.pal(11 , "Spectral" )))
gene.cp=c('lightgrey',rev(brewer.pal(11 , "Spectral" )))
clust.cp.separate = unique (c(cols25(25),alphabet2(26),glasbey(32),alphabet(26)))
clust.cp.graded = unique(c(stepped3(16),stepped(20),stepped2(20)))
```

Raw data from 10x Genomics cellranger software was imported and filtered to include only cells with >250 genes and >750 UMIS

Individual libraries were then manually inspected and upper and lower filters were selected for each prior to merging all data together using the Seurat 'integrate data' workflow based on anchor-finding with rpca algorythm.
# Analysis sections
```{r}
life.history = T #turn this off unless re-running analyses
setwd("/scratch/molevo/agcole/R/Aurelia_51k")
```

load data
```{r, echo=F}
#* from *GenerateAaAllData.AAUR2local.R*
load(file='AaAlldata.integrated.updatenames.Robj')
data1=SetIdent(Aa.Alldata,value='orig.ident')

```

# Life history comparisons
 
```{r, eval=F,echo=FALSE}
#if re-run, this puts the raw data libraries back into order.
#* otherwise this is saved
    data1$orig.ident = as.factor(data1$orig.ident)
    lib.names.original=dput(levels((data1$orig.ident)))
    order = c(7,9,8,10,1:6)
    data1@meta.data$orig.ident = factor(data1@meta.data$orig.ident,
                                        levels(data1@meta.data$orig.ident)[order])
    levels(data1@meta.data$orig.ident)
    Aa.Alldata = data1
```

```{r}
if (life.history)
{  
data1=  SetIdent(Aa.Alldata,value = 'orig.ident')
  polyp.all=WhichCells(data1,idents = levels(data1)[1:3])
  medusa.all = WhichCells(data1,idents = levels(data1)[6:10])
  strobila.all=WhichCells(data1,idents = levels(data1)[4])
  ephyra.all = WhichCells(data1,idents = levels(data1)[5])
  data1$lifehistory.tissue= data1$orig.ident
  data1$lifehistory.tissue[polyp.all]='polyp'
  data1$lifehistory.tissue[strobila.all]='strobila'
  data1$lifehistory.tissue[ephyra.all]='ephyra'

  Aa.Alldata$lifehistory.tissue = data1$lifehistory.tissue 
  
  Aa.Alldata$lifehistory.tissue = as.factor(Aa.Alldata$lifehistory.tissue)
  #re-order the libraries:
  Aa.Alldata$lifehistory.tissue=factor(Aa.Alldata$lifehistory.tissue,levels(Aa.Alldata$lifehistory.tissue)[c(7,8,1:6)])
  }
```

## Polyp vs medusa

```{r}
if (life.history)
{  
  polyp.all=WhichCells(data1,idents = levels(data1)[1:3])
  medusa.all = WhichCells(data1,idents = levels(data1)[6:10])
  strobila.all=WhichCells(data1,idents = levels(data1)[4])
  ephyra.all = WhichCells(data1,idents = levels(data1)[5])
  data1$lifehistory='aurelia'
  data1$lifehistory[polyp.all]='polyp'
  data1$lifehistory[strobila.all]='strobila'
  data1$lifehistory[medusa.all]='medusa'
  data1$lifehistory[ephyra.all]='ephyra'
  
  data1=SetIdent(data1,value='lifehistory')
  Aa.Alldata$lifehistory = data1@active.ident 
  }
```

```{r, echo=F,eval=F}
if (life.history)
{
data.temp=SetIdent(Aa.Alldata,value = 'lifehistory')
data1=subset(data.temp,cells = c(polyp.all,medusa.all,ephyra.all))
  polyp=WhichCells(data1,idents = levels(data1)[1])
  medusa = WhichCells(data1,idents = levels(data1)[2:3])
  polyp.genes=NULL
  polyp.genes = Aa_genes$gene_short_name[which(rowSums(data1@assays$RNA@counts[,polyp])>=1)] #31103 c | 29885 d
  medusa.genes=NULL
  medusa.genes = Aa_genes$gene_short_name[which(rowSums(data1@assays$RNA@counts[,medusa])>=1)] #40555 | 37903 d
  
  total.counts=NULL  
    mean.total.counts = NULL
  for (l in 1:2)
  {total.counts[[l]]=colSums(data1@assays$RNA@data[,WhichCells(data1,idents=levels(data1)[l])])
  mean.total.counts[l]=mean(total.counts[[l]])}
  a1=(polyp.genes)
  a2=(medusa.genes)
     num.gene=c(length(a1),length(a2))
        ST.venn.plot <- VennDiagram::venn.diagram(list(a1,a2),filename = NULL,
                                 category = c('polyp','medusa'),
                                 fill=LibCP[c(2,10)],#margin=1,
                                 cex=2,print.mode = 'raw',sigdigs=2,disable.logging=F)

        goi.polyp.vs.medusa=VennDiagram::calculate.overlap(list(a1,a2))
plot.new()
barplot(num.gene,col=LibCP[c(2,10)])+title('Polyp Vs Medusa | 1 or more reads')
}
```

```{r, warning=F,echo=F,eval=F}
if (life.history)
{

polyp.specific = setdiff(goi.polyp.vs.medusa$a1,goi.polyp.vs.medusa$a2)
medusa.specific = setdiff(goi.polyp.vs.medusa$a2,goi.polyp.vs.medusa$a1)
DotPlot(Aa.Alldata,'RNA',c(polyp.specific,medusa.specific),cols=magma(3),group.by = 'orig.ident',dot.min = 0.001,scale.by = 'size')+
  RotatedAxis()+labs(title='Polyp vs Medusa specific genes')+FontSize(0,)
}
```
### replicates: downsampling cells and genes
**Can we account for sequencing depth?**
This is 100 replicates sampling 1000 cells per stage with all genes >3 reads
```{r, eval=T, echo=F}
# if (life.history)

data.temp=SetIdent(Aa.Alldata,value = 'lifehistory')

    #should maybe remove non-expressed genes:
  keep = which(rowSums(data.temp@assays$RNA@counts) >= 10)
  data.temp=subset(data.temp,features = keep) #memory issue here.
  genes=Aa_genes[keep,]
  nr=100 #number of replicates
  replicates = NULL
  coi=NULL
  for (r in 1:nr)
  {
    coi[[r]]=WhichCells(data.temp,
                   downsample = 1000,seed = NULL)
  }
  
   goi=NULL
  for (r in 1:nr)#100)
  {
    goi[[r]]=c(1:length(genes$gene_short_name))
  }
  goi.lifehistory = NULL
  num.gene.all.rep = matrix (0L,nr,4)#4
  for (r in 1:nr)
  {
  data1=subset(data.temp,cells=coi[[r]])
  polyp=WhichCells(data1,idents = levels(data1)[1])
  medusa = WhichCells(data1,idents = levels(data1)[4])
  strobila=WhichCells(data1,idents = levels(data1)[2])
  #maybe ignore the strobila...
  ephyra = WhichCells(data1,idents = levels(data1)[3])
  
  polyp.genes=NULL
  polyp.genes = genes$gene_short_name[goi[[r]][which(rowSums(data1@assays$RNA@counts[,polyp])>=3)]] #30982
  medusa.genes=NULL
  medusa.genes = genes$gene_short_name[goi[[r]][which(rowSums(data1@assays$RNA@counts[,medusa])>=3)]] #40499
  strobila.genes = NULL
  strobila.genes = genes$gene_short_name[goi[[r]][which(rowSums(data1@assays$RNA@counts[,strobila])>=3)]] #38958
  ephyra.genes = NULL
  ephyra.genes = genes$gene_short_name[goi[[r]][which(rowSums(data1@assays$RNA@counts[,ephyra])>=3)]] #38958
  a1=(polyp.genes)
  a3=(strobila.genes)
  a4=(ephyra.genes)
  a2=(medusa.genes)
  num.gene.all.rep[r,]=c(length(a1),length(a3),length(a4),length(a2))
        ST.venn.plot <- VennDiagram::venn.diagram(list(a1,a2,a3,a4),filename = NULL,
                                 category = c('polyp','medusa','strobila','ephyra'),
                                 fill=LibCP[c(2,8,4,5)],#margin=1,
                                 cex=2,print.mode = 'percent',sigdigs=2,disable.logging=F)
  colnames(num.gene.all.rep) =c('polyp','strobila','ephyra','medusa')
# barplot(num.gene.all.rep)
    replicates[[r]] = ST.venn.plot
        goi.lifehistory[[r]]=VennDiagram::calculate.overlap(list(a1,a2,a3,a4))
    save(replicates,num.gene.all.rep,file='lifehistory.replicates.RData')

  }
```

#Fig1B
```{r, eval=T, echo=F, fig.height=4, fig.width=4}
 
      load(file='lifehistory.replicates.RData')
pdf('Fig1b.1.pdf')
      boxplot(num.gene.all.rep,col=LibCP[c(2,4,5,8)])#+title('Number of Expressed Genes')
      dev.off()
            boxplot(num.gene.all.rep,col=LibCP[c(2,4,5,8)])#+title('Number of Expressed Genes')
                  rm(replicates,num.gene.all.rep)
```

Run again without gene or cell filtering to get the raw numbers
#Fig1B.inset
```{r}
  # if (life.history)
{
  data1=SetIdent(Aa.Alldata,value = 'lifehistory')
  polyp=WhichCells(data1,idents = levels(data1)[1])
  medusa = WhichCells(data1,idents = levels(data1)[4])
  strobila=WhichCells(data1,idents = levels(data1)[2])
  #maybe ignore the strobila...
  ephyra = WhichCells(data1,idents = levels(data1)[3])
  
  polyp.genes=NULL
  polyp.genes = Aa_genes$gene_short_name[which(rowSums(data1@assays$RNA@counts[,polyp])>=1)] #31103 c | 29885 d
  medusa.genes=NULL
  medusa.genes = Aa_genes$gene_short_name[which(rowSums(data1@assays$RNA@counts[,medusa])>=1)] #40555 | 37903 d
  strobila.genes = NULL
  strobila.genes = Aa_genes$gene_short_name[which(rowSums(data1@assays$RNA@counts[,strobila])>=1)] #34476 | 31176 d
  ephyra.genes = NULL
  ephyra.genes = Aa_genes$gene_short_name[which(rowSums(data1@assays$RNA@counts[,ephyra])>=1)] #36893 | 33921 d
  total.counts=NULL  
    mean.total.counts = NULL
    num.gene.all = matrix (0L,1,4)#4
  for (l in 1:4)
  {total.counts[[l]]=colSums(data1@assays$RNA@data[,WhichCells(data1,idents=levels(data1)[l])])
  mean.total.counts[l]=mean(total.counts[[l]])}
  a1=(polyp.genes)
  a3=(strobila.genes)
  a4=(ephyra.genes)
  a2=(medusa.genes)
     num.gene.all=c(length(a1),length(a3),length(a4),length(a2))
        ST.venn.plot <- VennDiagram::venn.diagram(list(a1,a2,a3,a4),filename = NULL,
                                 category = c('polyp','medusa','strobila','ephyra'),
                                 fill=LibCP[c(2,10,4,8)],#margin=1,
                                 cex=2,print.mode = 'raw',sigdigs=2,disable.logging=F)
        goi.lifehistory.all=VennDiagram::calculate.overlap(list(a1,a2,a3,a4))
pdf('Fig1Bi.pdf',width = 3,height = 5)
barplot(num.gene.all,col=LibCP[c(2,4,5,8)])+title('genes with 1 or more reads')
dev.off()
        plot.new()
barplot(num.gene.all,col=LibCP[c(2,4,5,8)])+title('genes with 1 or more reads')
}
```

## Differentially Expressed Genes

```{r}
   if (life.history)
{
     data1=SetIdent(Aa.Alldata,value='lifehistory')
     data1@active.assay='RNA'
life.history.markers=FindAllMarkers(data1,
                                      logfc.threshold = 0.6,
                                      return.thresh = 0.0001,
                                      min.pct = 0.3,
                                      only.pos = T, 
                                      # max.cells.per.ident = 3000,
                                      verbose=F)
  
  life.history.markers[,8:12] = 'NA'
  names(life.history.markers)[8:12]=names(Aa_genes)[c(1,3:6)]
  # add GO terms associated with this list:
  for (i in 1:length(levels(data1@active.ident))) # 
  {
    x=life.history.markers[as.numeric(life.history.markers$cluster)==i,][1:length(which(as.numeric(life.history.markers$cluster)==i)),7]
    x2=Aa_genes[match(x,Aa_genes$gene_short_name),c(1,3:6)]
    life.history.markers[as.numeric(life.history.markers$cluster)==i,][1:length(which(as.numeric(life.history.markers$cluster)==i)),8:12]<-x2
  }
   
    save(life.history.markers,file='markers.lifehistory.RObj')
   }
```

#Fig1C

```{r, warning=FALSE, fig.width=24,fig.height=4}
load(file='/scratch/molevo/agcole/R/Aurelia_51k/markers.lifehistory.RObj')
View(life.history.markers)
DotPlot(data1,'RNA',(unique(life.history.markers$gene)),cols = c('lightgrey','darkred'),scale.by = 'radius',col.min = 0)&
    RotatedAxis()+FontSize(5,)

```


##### topGO | MF
```{r, echo=FALSE,}
   # if (life.history)
{
     library(topGO)

  Unfold <- Aa_annotations %>% 
    dplyr::mutate(`GOs` = strsplit(as.character(`GOs`), ",")) %>% 
    tidyr::unnest(`GOs`) 
  geneID2GO <- Unfold %>% split(x = .$`GOs`, f = .$gene_short_name)
  geneNames <- names(geneID2GO)
  #set dataset:
      all.markers =life.history.markers
  #process data
  GOdata = NULL
  myInterestingGenes = NULL
  geneList = NULL
  # this takes a really long time:
  for(i in 1:length(unique(all.markers$cluster)))
  {
    #identify the GOI from the DE list
    myInterestingGenes[[i]] <- all.markers$gene[all.markers$cluster==unique(all.markers$cluster)[i]] #list of genes you want to perform GO enrichment for
    geneList[[i]] <- factor(as.integer(geneNames %in% myInterestingGenes[[i]]))
    names(geneList[[i]]) <- geneNames
    GOdata[[i]] <- new("topGOdata",ontology = "BP", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
  }
   }
```


```{r, echo=FALSE,}
   # if (life.history)
{
     # filter and generate figures.
  p=NULL
  resultFis=NULL
  topGO.lifehistory=NULL
  for(i in 1:length(unique(all.markers$cluster)))
  {
    #run the test...
    resultFis[[i]] <- runTest(GOdata[[i]], algorithm = "classic", statistic = "fisher") 
    pvalFis <- score(resultFis[[i]])
    #filter for only >0.05
    pvalFis = pvalFis[pvalFis>=0.05]
    allRes_intgenes<- GenTable(GOdata[[i]], pvalues = resultFis[[i]], orderBy = "pvalues", topNodes=30)
    allRes_intgenes$pvalues<-as.numeric(allRes_intgenes$pvalues)
    #convert NA values to zero; only if p value is so small it cannot be displayed in r
    allRes_intgenes[is.na(allRes_intgenes)]<-0.00000001
    
    #plot GOenrichment 
    
    david_col=LibCP[c(2,10,4,8)]
    topGO.lifehistory[[i]]=
      ggplot2::ggplot(allRes_intgenes, ggplot2::aes(x=(reorder(Term,(-log10(pvalues)))), y=(-log10(pvalues)))) +
      ggplot2::stat_summary(geom = "bar", fun = mean, position = "dodge",
                            col=david_col[i],fill=david_col[i]) +
      ggplot2::coord_flip()+
      ggplot2::xlab("Biological Process") +
      ggplot2::ylab("Enrichment -log10 p-value") +
      ggplot2::labs(title="GO enrichment",subtitle=levels(all.markers$cluster)[i])
      
    print(topGO.lifehistory[[i]])
  }

  save(topGO.lifehistory,file='lifehistory.topGO.RObj')
   }
```

evaluate the topGO results:
```{r, echo=FALSE, eval=T}
load(file='lifehistory.topGO.RObj')
print(topGO.lifehistory[[1]]+topGO.lifehistory[[2]]+topGO.lifehistory[[3]]+topGO.lifehistory[[4]]+plot_layout(ncol=2))
```
evaluate the topGO results:
#Fig1D
```{r, echo=FALSE, eval=T, fig.height=8,fig.width=8}
pdf('Fig1E.pdf',height = 8,width = 8)
print(topGO.lifehistory[[1]]+topGO.lifehistory[[4]])
dev.off()
print(topGO.lifehistory[[1]]+topGO.lifehistory[[4]])
# rm(topGO.lifehistory)
```

```{r}
data1<-SetIdent(Aa.Alldata,value='lifehistory')
data1@active.assay='RNA'
      all.markers.TF <- FindAllMarkers(data1,
                                     logfc.threshold = 0.1,
                                     features = intersect(AaTF_list,rownames(data1)),
                                     min.pct = 0.001,
                                     only.pos = TRUE, 
                                     max.cells.per.ident = 3000,
                                     return.thresh = 0.0001,
                                  verbose = F)
    
    all.markers.TF[,8:12] = 'NA'
    names(all.markers.TF)[8:12]=names(genes)[c(1,3:6)]
    # add GO terms associated with this list:
    for (i in 1:length(levels(data1@active.ident))) # 
    {
      x=all.markers.TF[as.numeric(all.markers.TF$cluster)==i,][1:length(which(as.numeric(all.markers.TF$cluster)==i)),7]
      x2=genes[match(x,genes$gene_short_name),c(1,3:6)]
      all.markers.TF[as.numeric(all.markers.TF$cluster)==i,][1:length(which(as.numeric(all.markers.TF$cluster)==i)),8:12]<-x2
    }  
    
    list_TF = NULL
    for (i in 1:length(levels(data1@active.ident)))
    {
      x=all.markers.TF[as.numeric(all.markers.TF$cluster)==i,][1:min(20,length(which(as.numeric(all.markers.TF$cluster)==i))),7]
        list_TF=c(list_TF,x)
    }
    list_TF=unique(list_TF)
    TF.plot=DotPlot(data1, features = unique(list_TF), 
                    scale.by='size',# group.by = 'ID.separate', 
                    col.min = 0, col.max = 3,  
                    cols = c('lightgrey','darkred')) + 
      RotatedAxis() +FontSize(6,6) +NoLegend()+#coord_flip()+
      labs(title = 'Top 20 TFs',subtitle = 'p-val < 0.001')
  
print(TF.plot)
life.history.DETF=all.markers.TF
save(life.history.DETF,file='life.history.DETF.RObj')
write.csv(life.history.DETF,file='S1.3_lifecycleDETF.csv')
```

```{r}
life.history.DETF=life.history.DETF[life.history.DETF$pct.2<=0.2,]
  goi = NULL
    for (i in 1:length(levels(data1@active.ident)))
    {
      x=life.history.DETF[as.numeric(life.history.DETF$cluster)==i,][1:min(20,length(which(as.numeric(life.history.DETF$cluster)==i))),7]
      x=x[]
        goi=c(goi,x)
    }
    goi=unique(goi)

    DotPlot(data1,'RNA', features = goi, 
                    scale.by='size', split.by = 'lifehistory', 
                    col.min = 0, col.max = 3,  
                    cols =LibCP[c(2,4,5,10)]) + 
      RotatedAxis() +FontSize(6,6) +theme(legend.position = 'bottom')
        goi=goi[c(2,3,11,14,15,30,33,35,36,44,52,56,57,58,59,64)]#
    pdf('fig1F.vln.pdf',height = 8,width = 8)
      VlnPlot(data1,goi)
      dev.off()
VlnPlot(data1,goi)
      
```

```{r}
Aa.Alldata=data1
save(Aa.Alldata,file='Aa.Alldata.newnames.RObj')
```

