---
title: "Aurelia Notebook | AAUR2 mapping"
output: html_notebook
---

#Aurelia Cell Atlas | Complete

```{r, echo=F, warning=FALSE}
 ## Where you want to work
#load libraries: 
library(Seurat,quietly=T)
library(RColorBrewer,quietly=T)
library(patchwork,quietly=T)
library(ggplot2,quietly=T)
library(pals,quietly=T)
library(readxl,quietly=T)
library(SeuratWrappers,quietly=T)
library(tidyr,quietly=T)

```

```{r colours, echo=F}

load (file="Aaur2.newnames.RData")

```

Raw data from 10x Genomics cellranger software was imported and filtered to include only cells with >250 genes and >750 UMIS

Individual libraries were then manually inspected and upper and lower filters were selected for each prior to merging all data together using the Seurat 'integrate data' workflow based on anchor-finding with rpca algorythm.
# Analysis sections
```{r}
all.data = T
#use this to turn analyses on and off
```

# AllData integrated dataset: reductions and clustering
 **SuperClusters:**
```{r}
if (!exists('Aa.Alldata'))
load(file='Aa.Alldata.newnames.RObj')

data1=SetIdent(Aa.Alldata,value='orig.ident')
data1@active.assay='integrated'
data1$orig.ident=as.factor(data1$orig.ident)
data1$orig.ident=factor(data1$orig.ident,levels(data1$orig.ident)[c(7,9,8,10,1:6)])
levels(data1)
Aa.Alldata=data1
```

#### PCAs
```{r}
if(all.data)
{#saved

    data1 <- ScaleData(data1, verbose = FALSE)
    data1 <- RunPCA(data1,
                    features = data1@assays$integrated@var.features,
                    npcs = 50, verbose = FALSE)
    PCAPlot(data1)
    }
```
#### Dimensions
```{r}
if(all.data)
{
  # Determine percent of variation associated with each PC
pct <- data1[["pca"]]@stdev / sum(data1[["pca"]]@stdev) * 100

# Calculate cumulative percents for each PC
cumu <- cumsum(pct)

# Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
co1 <- which(cumu > 90 & pct < 5)[1]
# Determine the difference between variation of PC and subsequent PC
co2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1

# choose the minimum of these two metrics as the PCs covering the majority of the variation in the data
pcs <- min(co1, co2)
d = 1:pcs
print(pcs)
}
```
#### UMAP
```{r, warning=FALSE, }
if(all.data)
{
  data1 <- RunUMAP(data1, reduction = "pca", 
                     dims = d,
                     n.neighbors = 20L, #how many similar cells do you expect save==10
                     spread =0.6, #how big is the axis space: 0.2 for lineages; 1 for separation
                     min.dist = 0.2, #how close to plot each cell higher=spread
                     local.connectivity = 100,#overall how connected is the graph save==1
                     reduction.name = 'umap.int',reduction.key = 'umap.int',
                     verbose = F,
                   seed.use = 10)
}
```

```{r}
DimPlot(data1,group.by = 'orig.ident',cols=alpha(LibCP,1))&NoAxes()+NoLegend()&labs(title='medusa top')&
DimPlot(data1,group.by = 'orig.ident',cols=alpha(rev(LibCP),1),
        order=levels(as.factor(data1$orig.ident)))+labs(title='polyp top')&NoAxes()

```
Plotted with both layering; all libraries are presented throughout all clusters. This indicates that the integration was successful. It is also evident here that 

A) digestive gland cells are **NOT** in the subumbrella (purple)
B) polyp cells (blue) are **NOT** in the striated muscle (ID shown later)
 
#### Neighbours
```{r, }
if(all.data)
{
    data1 <- FindNeighbors(data1, 
                           reduction = "pca", 
                           dims = d,
                           annoy.metric = 'cosine',
                           verbose=F,
                           k.param = 10) 
}
```
#### Clusters
```{r, }
if(all.data)
{
  data1 <- FindClusters(data1, verbose=F,
                      resolution = 0.1) 
data1 <- BuildClusterTree(data1,
                          reorder = TRUE,
                          reorder.numeric = T,verbose = F)
data1$ID.cluster=data1@active.ident
}
```

```{r, eval=FALSE}
DimPlot(data1, reduction = "umap.int", label = TRUE, repel = TRUE,cols=clust.cp.separate)&NoAxes()

```

Neuroglandular clusters appear to emerge from the **inner gastrodermal epithelia**, *excluding the striated muscle.*

```{r}
FeaturePlot(data1,c('nFeature_RNA','nCount_RNA'),order=T,cols=gene.cp)&NoAxes()&NoLegend()
  # VlnPlot(data1,'nFeature_RNA')
```
**BUT** this connection is largely low-information cells. *So is this biological or technical? This 'bridge' shifted with stricter filtering.* Was from center or even inner cell layer with more 'noise' included. Note that the separation between inner and outer cell layers is somewhat arbitrary (clustering) and indeed there are 'inner cell layer' cells in the 'outer cell layer' supercluster. Moved these later in the subsets.

#### assign names
```{r }
if(all.data)
{
data1@active.assay = 'RNA'
  data1<-SetIdent(data1,value = 'ID.cluster')
  print(DimPlot(data1,cols=clust.cp.separate,label=T))

   clusterNames<- read_excel('DataS3.xlsx',
                            sheet = 'AGC.markers.integrated')
                            # sheet='AGC.markers.all')
  # 'In situ genes'
  
  goi = Aa_genes$gene_short_name[match(clusterNames$marker,Aa_genes$name)]
  #assign cluster ID to the individual libraries
  data1<-ScaleData(data1,features = goi,verbose=F)#, split.by = 'orig.ident') #works with scaled gene values, so must scale all of these first (in case they are not within the VarFeatures set!)
     clName = vector()
    
    #generate a matrix of values of each cluster for each gene:
    m=AverageExpression(data1,features=goi,slot='scale.data') #Seurat has a function for this
    
    for (j in 1:length(levels(data1@active.ident))) #for each cluster set
    {
      clName[j]=as.integer(which.max(m$RNA[,j])) #choose the highest value
    }
    sort(clName) #this prints which IDs were selected; sanity check that it did what you wanted... should have no unwanted replicated IDs
    clusterNames$ID[clName]
    #Note: here we merge the cnidocytes, central, and dig.gland clusters
  
  #use the wanted order from the spreadsheet to re-order the clusters:
  #first order the identities..
  data1@active.ident = factor(data1@active.ident,
                              levels(data1@active.ident)[order(clName)])
  #set the names to your IDs; once the gene list works, it is really easy to then update your cluster names with your excel spreadsheet.
  levels(data1@active.ident) = clusterNames$ID[clName][order(clName)]
  #save the IDs in metadata:
  data1@meta.data$IDs = data1@active.ident
data1@active.assay = 'integrated'
  print(DimPlot(data1,cols=clust.cp.separate,label=T))

  Aa.Alldata=data1
}
```

```{r, echo=FALSE}
data1=Aa.Alldata
data1@active.assay = 'integrated'
  clust.cp =c(clust.cp.separate[c(1,8,7,6,2,3,5,4,9)])
  {
        ids.cluster.sample = as.data.frame(table(Idents(data1), data1@meta.data$lifehistory.tissue))
    colnames(ids.cluster.sample) = c('ID', 'sample', 'CellCount')
    
    #generate your sample and cluster UMAP plots:
    sample.plot =      DimPlot(
      data1,
      group.by = 'orig.ident',
      order = rev(levels(data1$orig.ident)),
      cols = rev(LibCP)
    ) + NoAxes() +
      labs(title = 'Time | sample origin')#+NoLegend()
    
    cluster.plot = DimPlot(
      data1,
      label = F,
      label.size = 4,
      repel = T,
      #order=rev(levels(data1@active.ident)),
      #ordering is optional
      cols = clust.cp
    ) + NoAxes() +
      labs(title = 'Clusters | ID') + NoLegend()
    cluster.plot
    #barplot of cluster identities in each sample:
    
    dist.clust2 =
      ggplot(ids.cluster.sample, aes(fill = ID, y = CellCount,
                                     x = sample)) +
      geom_bar(
        mapping = aes(
          fill = ID,
          y = (CellCount),
          x = (sample)
        ),
        position = "fill",
        stat = "identity",
        width = 0.5
      ) +
      scale_fill_manual(values = clust.cp) +
      theme(axis.text.x = element_text(
        #face="bold", color="#993333",
        size = 8,
        angle = -45,
        hjust = 0,
        vjust = 0.5
      )) +
      geom_area(
        mapping = aes(
          fill = ID,
          y = (CellCount),
          x = as.integer(sample)
        ),
        position = "fill",
        stat = "identity",
        alpha = 0.3 ,
        size = .5,
        colour = "white"
      ) +
      geom_bar(
        mapping = aes(
          fill = ID,
          y = (CellCount),
          #this re-plots the bars over the area
          x = (sample)
        ),
        position = "fill",
        stat = "identity",
        width = 0.5
      ) +
      ggtitle("Distribution of cell types in time and space")+theme(legend.position = 'bottom')
    dist.clust2
    
    data1<-SetIdent(Aa.Alldata,value='ID.separate')
            ids.cluster.sample = as.data.frame(table(Idents(data1), data1@meta.data$lifehistory.tissue))
    colnames(ids.cluster.sample) = c('ID', 'sample', 'CellCount')
  }
  #visualize your plots; can also do these individually as desired
  print(cluster.plot+dist.clust2) 
```

# Differential Expression
##### All genes
**Set of all up-regulated genes** generated with: 

- 200 cell downsampling 
- logfc >= 0.6 
- at least 30% of cells expressing the gene
- p value <= 0.0001
  
```{r,warning=FALSE}

  data1=SetIdent(Aa.Alldata,value='IDs')
{
  data1@active.assay='RNA'
  genes=Aa_genes
    all.markers <- FindAllMarkers(data1,
                                  logfc.threshold = 0.6,
                                  return.thresh = 0.0001,
                                  min.pct = 0.3,
                                  only.pos = TRUE, 
                                  max.cells.per.ident = 200,
                                  verbose = F)

      all.markers[,8:12] = 'NA'
      names(all.markers)[8:12]=names(genes)[c(1,3:6)]
      # add GO terms associated with this list:
      for (i in 1:length(levels(data1@active.ident))) # 
      {
        x=all.markers[as.numeric(all.markers$cluster)==i,][1:length(which(as.numeric(all.markers$cluster)==i)),7]
        x2=genes[match(x,genes$gene_short_name),c(1,3:6)]
        all.markers[as.numeric(all.markers$cluster)==i,][1:length(which(as.numeric(all.markers$cluster)==i)),8:12]<-x2
      }  

    
    #generate a collated list of unique DE genes
    list = NULL
    for (i in  1:length(levels(data1@active.ident)))
    {
      x=all.markers[as.numeric(all.markers$cluster)==i,][1:min(5,length(which(as.numeric(all.markers$cluster)==i))),7]
        list=c(list,x)
    }
    
    #Image the list
    DEG.plot=DotPlot(data1,'RNA', features = unique(c(list)), 
                     scale.by='radius' , col.min = 0, col.max = 3,  
                     cols = c('lightgrey','darkred')) + 
      RotatedAxis() +FontSize(6,8) +#coord_flip()+
      # NoLegend()+
      labs(title = 'Top 5 DEGs',subtitle = 'p-val < 0.001')

}
    all.markers.Alldata=all.markers 
  save(all.markers.Alldata,file = 'AllData.integrated2.markers.RObj')
```
    
```{r,fig.height=6,fig.width=12}
load(file = 'AllData.integrated2.markers.RObj') 
  data1=SetIdent(Aa.Alldata,value='IDs')
list = NULL
    for (i in  1:length(levels(data1@active.ident)))
    {
      x=all.markers.Alldata[as.numeric(all.markers.Alldata$cluster)==i,][1:min(5,length(which(as.numeric(all.markers.Alldata$cluster)==i))),7]
        list=c(list,x)
    }
    
    #Image the list
    DEG.plot=DotPlot(data1,'RNA', features = unique(c(list)), 
                     scale.by='radius' , col.min = 0, col.max = 3,  
                     cols = c('lightgrey','darkred'),scale=T) + 
      RotatedAxis() +FontSize(6,8) +#coord_flip()+
      # NoLegend()+
      labs(title = 'Top 5 DEGs',subtitle = 'p-val < 0.001')
all.markers.Alldata
  DimPlot(data1,cols=clust.cp.separate[c(1,8,7,6,2,3,5,4,9)])+NoAxes()+plot_layout(ncol = 2)&DEG.plot
 
```
#Fig2A
```{r,fig.width=3,fig.height=3}
pdf('fig2A.pdf',width = 3,height = 3)
  DimPlot(Aa.Alldata,cols=clust.cp.separate[c(1,8,7,6,2,3,5,4,9)])+NoAxes()+NoLegend()
dev.off()
  DimPlot(Aa.Alldata,cols=clust.cp.separate[c(1,8,7,6,2,3,5,4,9)])+NoAxes()+NoLegend()
```

```{r,warning=FALSE}

  data1=SetIdent(Aa.Alldata,value='IDs')
{
  data1@active.assay='RNA'
    all.markers <- FindAllMarkers(data1,features = AaTF_list,
                                  return.thresh = 0.001,
                                  min.pct = 0.01,
                                  only.pos = TRUE, 
                                  verbose = F)

      all.markers[,8:12] = 'NA'
      names(all.markers)[8:12]=names(genes)[c(1,3:6)]
      # add GO terms associated with this list:
      for (i in 1:length(levels(data1@active.ident))) # 
      {
        x=all.markers[as.numeric(all.markers$cluster)==i,][1:length(which(as.numeric(all.markers$cluster)==i)),7]
        x2=genes[match(x,genes$gene_short_name),c(1,3:6)]
        all.markers[as.numeric(all.markers$cluster)==i,][1:length(which(as.numeric(all.markers$cluster)==i)),8:12]<-x2
      }  
    all.markers.Alldata.TF=all.markers 
  save(all.markers.Alldata.TF,file = 'AllData.integrated2.TFmarkers.RObj')
  }
```

```{r,warning=FALSE} 
  data1=SetIdent(Aa.Alldata,value='IDs')
load(file = 'AllData.integrated2.TFmarkers.RObj')
all.markers=all.markers.Alldata.TF
    #generate a collated list of unique DE genes
    list = NULL
    for (i in  1:length(levels(data1@active.ident)))
    {
      x=all.markers[as.numeric(all.markers$cluster)==i,][1:min(5,length(which(as.numeric(all.markers$cluster)==i))),7]
        list=c(list,x)
    }
    
    #Image the list
    DEG.plot=DotPlot(data1,'RNA', features = unique(c(list)), 
                     scale.by='radius' , col.min = 0, col.max = 3,  
                     cols = c('lightgrey','darkred')) + 
      RotatedAxis() +FontSize(6,8) +#coord_flip()+
      # NoLegend()+
      labs(title = 'Top 5 DEGs',subtitle = 'p-val < 0.001')

print(DEG.plot)

```

#Fig2C
```{r}
import.alra = F #once saved
if (import.alra)
{#import the alra assay into the new alldata object: currently data1
load (file='AaAlldata.integrated.AAUR2.Robj')
data1@assays$alra = Aa.Alldata@assays$alra
data1@assays$alra@data@Dimnames[[1]]=Aa_genes$gene_short_name
Aa.Alldata=data1}
data1=SetIdent(Aa.Alldata,value = 'IDs')
data1@active.assay='alra'
list=c('AastMHC5','AAUR2.48938','APLP-like-3','FOXA2-like-1','AaMUC2-like','AaTBA1-like3','FOS-like-1','AAUR2.29811','PRD13-like-1')
list=Aa_genes$gene_short_name[match(list,Aa_genes$name)]
pdf('Fig.2C.pdf',width=14,height = 2)
print(FeaturePlot(data1,list,order=T,cols=gene.cp,ncol = 9)&NoAxes()&NoLegend())
dev.off()
print(FeaturePlot(data1,list,order=T,cols=gene.cp,ncol = 9)&NoAxes()&NoLegend())

```
##### with DNA binding domains
**Set of all up-regulated *DNA-binding* genes** generated with: 

- 200 cell downsampling 
- logfc >= 0.2 (default) 
- at least 1% of cells expressing the gene: this accounts for low expression of this class of genes
- p value <= 0.001
  
```{r}
{
  data1=SetIdent(Aa.Alldata,value='IDs')
  data1@active.assay='RNA'
  all.markers.TF <- FindAllMarkers(data1,
                                     # logfc.threshold = 0.4,
                                     features = intersect(AaTF_list,rownames(data1)),
                                     min.pct = 0.01,
                                     only.pos = TRUE, 
                                     max.cells.per.ident = 1000,
                                     return.thresh = 0.001,
                                  verbose = F)
    
    all.markers.TF[,8:12] = 'NA'
    names(all.markers.TF)[8:12]=names(genes)[c(1,3:6)]
    # add GO terms associated with this list:
    for (i in 1:length(levels(data1@active.ident))) # 
    {
      x=all.markers.TF[as.numeric(all.markers.TF$cluster)==i,][1:length(which(as.numeric(all.markers.TF$cluster)==i)),7]
      x2=genes[match(x,genes$gene_short_name),c(1,3:6)]
      all.markers.TF[as.numeric(all.markers.TF$cluster)==i,][1:length(which(as.numeric(all.markers.TF$cluster)==i)),8:12]<-x2
    }  
    
    
   all.markers.TF.Alldata=all.markers.TF
      save(all.markers.TF.Alldata,file = 'AllData.integrated2.TF.markers.RObj')
      }

```
image results
```{r}
# load(file = 'AllData.integrated2.TF.markers.RObj')
  all.markers.TF.Alldata
  list_TF = NULL
    for (i in 1:length(levels(data1@active.ident)))
    {
      x=all.markers.TF.Alldata[as.numeric(all.markers.TF.Alldata$cluster)==i,][1:min(20,length(which(as.numeric(all.markers.TF.Alldata$cluster)==i))),7]
        list_TF=c(list_TF,x)
    }
    list_TF=unique(list_TF)
    TF.plot=DotPlot(data1, features = unique(list_TF), 
                    scale.by='radius', 
                    col.min = 0, col.max = 3,  
                    cols = c('lightgrey','darkred')) + 
      RotatedAxis() +FontSize(6,6) +NoLegend()+#coord_flip()+
      labs(title = 'Top 5 TFs',subtitle = 'p-val < 0.001')
   DimPlot(data1,cols=clust.cp.separate[c(1,8,7,6,2,3,5,4,9)])+NoAxes()+plot_layout(ncol = 2)& TF.plot

```

#save alldata
```{r, }
if(all.data)
{    
  Aa.Alldata=data1
    save(Aa.Alldata,file='Aa.Alldata.newnames.RObj')
}
```
