---
title: "Aurelia Notebook | AAUR2 mapping"
output:
  html_notebook: default
  pdf_document: default
---

#Aurelia Cell Atlas | Complete
Here we use the previous mapping and just re-run the analyses via cluster

```{r, echo=F, warning=FALSE}

#load libraries: 
library(Seurat,quietly=T)
library(RColorBrewer,quietly=T)
library(patchwork,quietly=T)
library(ggplot2,quietly=T)
library(pals,quietly=T)
library(monocle3,quietly=T)
library(readxl,quietly=T)
library(SeuratWrappers,quietly=T)
library(tidyr,quietly=T)

load (file="Aaur2.newnames.RData")
if(!exists('Aa.Alldata'))
  load (file='Aa.Alldata.newnames.RObj')
genes=Aa_genes
LibCP =brewer.paired(10)
gene.cp.dark=c('grey30',rev(brewer.pal(11 , "Spectral" )))
gene.cp=c('lightgrey',rev(brewer.pal(11 , "Spectral" )))
clust.cp.separate = unique (c(cols25(25),alphabet2(26),glasbey(32),alphabet(26)))
clust.cp.graded = unique(c(stepped3(16),stepped(20),stepped2(20)))

do.over=F
if (do.over)
{
Aa.Alldata<-SetIdent(Aa.Alldata, value='IDs')
data1.subsets <- SplitObject(Aa.Alldata)
order = match(levels(Aa.Alldata),names(data1.subsets))
data1.subsets=data1.subsets[order]

for (i in 1:length(names(data1.subsets)))
{
  data1=data1.subsets[[i]]
  data1$orig.ident=as.factor(data1$orig.ident)
data1$orig.ident=factor(data1$orig.ident,levels(data1$orig.ident)[c(7,9,8,10,1:6)])
levels(data1$orig.ident)
data1.subsets[[i]]=data1

}

}else{

  load(file = 'AllData.integrated.subsets.newnames.RObj')
  load(file = 'AllData.integrated.subsets.newnames.markers.RObj')
  load(file = 'AllData.integrated.subsets.newnames.TF.markers.RObj')
  
}

neur.cp=c("#33a02c",stepped2(8)[5:8],stepped(12)[10:12],'darkgreen',"#ff7f00",stepped(8)[5:8],stepped3(8)[7:8],'orange','orange3','darkorange4')

all.clusters.cp=c("#1F78C8","darkblue","#a6cee3",brewer.blues(10)[2:6],"#FFD700" ,'#f9ffa1','#f2db00','#e7a300','orange4',"#565656",brewer.greys(20)[c(4,8,10,12,20,14,16)],"#ff0000",brewer.reds(20)[c(4,6,8,10,15)],"#6A33C2", brewer.purples(15)[5:12],"#FB6496",brewer.rdpu(20)[5:11],neur.cp)
names(all.clusters.cp)=levels(SetIdent(Aa.Alldata,value='ID.separate'))

```


```{r}
#* to turn analyses on and off...
cluster.subsets = T
```

Raw data from 10x Genomics cellranger software was imported and filtered to include only cells with >250 genes and >750 UMIS

Individual libraries were then manually inspected and upper and lower filters were selected for each prior to merging all data together using the Seurat 'integrate data' workflow based on anchor-finding with rpca algorythm.

## Standard analysis
This runs all cell populations (coarse clustering of entire dataset) through a standard Seurat-based analysis pipeline and generates fine (res=1) and coarse (res=0.2) clustering for downstream analysis.
```{r SeuratPipeline,warning=FALSE, message=FALSE}
if(cluster.subsets)
{
  for (i in 1:length(names (data1.subsets)))
  {
    data1=data1.subsets[[i]]
    data1@active.assay = 'integrated'
    
    data1 <- FindVariableFeatures(data1,nfeatures = 2000,verbose = F)#

    # #use the full dataset scaling:
    Aa.Alldata@active.assay='integrated'
    coi = colnames(data1)
    t=ScaleData(Aa.Alldata,features = data1@assays$integrated@var.features)
    data1@assays$integrated@scale.data = t@assays$integrated@scale.data[,coi]
    rm(t)
    data1 <- RunPCA(data1, pcs.compute = 50,verbose = F) #this is now failing...
    ElbowPlot(object = data1, ndims = 50)
    # Determine percent of variation associated with each PC
    pct <- data1[["pca"]]@stdev / sum(data1[["pca"]]@stdev) * 100
    # Calculate cumulative percents for each PC
    cumu <- cumsum(pct)
    # Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
    co1 <- which(cumu > 90 & pct < 5)[1]
    # Determine the difference between variation of PC and subsequent PC: last point where change of % of variation is more than 0.1%
    co2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1
    # choose the minimum of these two metrics as the PCs covering the majority of the variation in the data
    pcs <- min(co1, co2)
    d = 1:pcs
    data1 <- RunUMAP(data1, reduction ="pca", n.neighbors = 10L,spread =1,seed.use = 0, 
                     dims = d,reduction.name ='umap',
                     reduction.key ='umap',min.dist = 0.3,
                     local.connectivity = 10,verbose = F)#
    data1 <- FindNeighbors(object = data1,reduction ="pca",dims = d,
                           nn.method = 'annoy',  annoy.metric = 'cosine',verbose = F)
    data1 <- FindClusters(object = data1,resolution = 1,random.seed = 0,verbose = F)
    # #look at relationship between clusters
    data1 <- BuildClusterTree(object = data1, reorder = TRUE,
                              reorder.numeric = TRUE,verbose = F)#, dims = c(1:d))
    data1$IDs.fine = data1@active.ident
    data1 <- FindClusters(object = data1,resolution = 0.2,random.seed = 0,verbose = F)
    # #look at relationship between clusters
    data1 <- BuildClusterTree(object = data1, reorder = TRUE,
                              reorder.numeric = TRUE,verbose = F)#, dims = c(1:d))
    data1$IDs.coarse = data1@active.ident
    data1@reductions=data1@reductions[c(1,3,2)]
    data1.subsets[[i]] = data1
  }#this sets coarse clustering as default
}
```

Plot the distribution summaries for all clusters. This will change depending upon whether or not subsequent analyses have been run!
```{r, eval=T, warning=F,message=F}

clust.cp=clust.cp.separate
for (i in 1:length(names (data1.subsets)))
{ 
  data1=data1.subsets[[i]]
  ids.cluster.library = as.data.frame(table(Idents(data1), data1@meta.data$orig.ident))
  colnames(ids.cluster.library) = c('ID','Library','CellCount')
  
  print(
    DimPlot(data1, label = F,label.size = 5, repel = T,
            cols = clust.cp)+NoLegend()+NoAxes()+
      labs(title = names(data1.subsets)[i],subtitle = i) +
      ggplot(ids.cluster.library, aes(fill=ID, y= (CellCount),
                                      x=Library)) +
      geom_bar(mapping =aes(fill=ID, y= (CellCount),
                            x=(Library)),
               position="fill", stat="identity", width = 0.5)+
      scale_fill_manual(values = clust.cp)+
      theme(axis.text.x = element_text(#face="bold", color="#993333", 
        size=8, angle=-45,hjust=0,vjust = 0.5))+
      geom_area(mapping =aes(fill=ID, y= (CellCount),
                             x=as.integer(Library)),
                position="fill", stat="identity",alpha=0.2 , size=.5, colour="white") +
      geom_bar(mapping =aes(fill=ID, y= (CellCount),#this re-plots the bars over the area
                            x=(Library)),
               position="fill", stat="identity", width = 0.5)+labs(title = names(data1.subsets)[i],subtitle = i)
  )
}
```

```{r}
if(cluster.subsets)
{
  save(data1.subsets,file = 'AllData.integrated.subsets.newnames.RObj')
}
```

#### Refine subsets ----
Cnidocytes and mature are separate clusters in the main dataset. Put these back together as a combined subset; also re-evaluated n1 and n2 clusters together below

# Outer layer
```{r outer_names, warning=FALSE}
if(cluster.subsets)
{
  data1=data1.subsets$outer.epidermis
  clusterNames<- read_excel('DataS3.xlsx',
                            sheet = 'OuterClusters')
  
  goi = Aa_genes$gene_short_name[match(clusterNames$marker,Aa_genes$name)]
  data1@active.assay='RNA'
  #assign cluster ID to the individual libraries
  data1<-ScaleData(data1,features = goi,verbose = F) 
  clName = vector()
  
  #generate a matrix of values of each cluster for each gene:
  m=AverageExpression(data1,features=goi,slot='scale.data') #Seurat has a function for this
  
  for (j in 1:length(levels(data1@active.ident))) #for each cluster set
  {
    clName[j]=as.integer(which.max(m$RNA[,j])) #choose the highest value
  }
  sort(clName)#for long lists you might want to know what is missing
  clusterNames$ID[clName]
  DimPlot(data1,cols=clust.cp.separate,label=T)&NoAxes()
  
  #use the wanted order from the spreadsheet to re-order the clusters:
  #first order the identities..
  data1@active.ident = factor(data1@active.ident,
                              levels(data1@active.ident)[order(clName)])
  #set the names to your IDs; once the gene list works, it is really easy to then update your cluster names with your excel spreadsheet.
  levels(data1@active.ident) = clusterNames$ID[clName][order(clName)]
  #save the IDs in metadata:
  data1@meta.data$IDs = data1@active.ident
  data1@active.assay='integrated'
  print(DimPlot(data1,group.by = c('IDs'),label=F,cols = (clust.cp.separate))+NoAxes())
  # data1@reductions=data1@reductions[c(1,3,2)]
  data1.subsets$outer.epidermis=data1
}
```

##### topGO | Outer epidermis
```{r outer_topGO, echo=FALSE, eval=F}
if (cluster.subsets)
{
data1=data1.subsets$outer.epidermis
all.markers <- FindAllMarkers(data1,
                              return.thresh = 0.01,
                              min.pct = 0.2,
                              only.pos = TRUE,
                              max.cells.per.ident = 100,
                              verbose = F)
library(topGO)

resultFis=NULL
topGO.outer.layer=NULL

Unfold <- Aa_annotations %>% 
  dplyr::mutate(`GOs` = strsplit(as.character(`GOs`), ",")) %>% 
  tidyr::unnest(`GOs`) 
geneID2GO <- Unfold %>% split(x = .$`GOs`, f = .$gene_short_name)
geneNames <- names(geneID2GO)
#set dataset:
all.markers
#process data
GOdata.BP.outer.layer = NULL
GOdata.MF.outer.layer = NULL
GOdata.CC.outer.layer = NULL
myInterestingGenes = NULL
allRes_intgenes = NULL

geneList = NULL
for (r in 1:3)
{
  # this takes a really long time:
  for(i in 1:length(unique(all.markers$cluster)))
  {
    #identify the GOI from the DE list
    myInterestingGenes[[i]] <- all.markers$gene[all.markers$cluster==unique(all.markers$cluster)[i]] #list of genes you want to perform GO enrichment for
    geneList[[i]] <- factor(as.integer(geneNames %in% myInterestingGenes[[i]]))
    names(geneList[[i]]) <- geneNames
    GOdata.BP.outer.layer[[i]] <- new("topGOdata",ontology = "BP", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.MF.outer.layer[[i]] <- new("topGOdata",ontology = "MF", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.CC.outer.layer[[i]] <- new("topGOdata",ontology = "CC", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    go.type=c("Biological Process",'Cellular Compartment','Molecular Function')
    
    if(r==1)
    {GOdata=GOdata.BP.outer.layer}else if(r==2)
    {GOdata=GOdata.CC.outer.layer}else if (r==3)
    {GOdata=GOdata.MF.outer.layer}  
    
    
    #run the test...
    resultFis[[i]] <- runTest(GOdata[[i]], algorithm = "classic", statistic = "fisher")
    pvalFis <- score(resultFis[[i]])
    #filter for only >0.05
    pvalFis = pvalFis[pvalFis>=0.05]
    allRes_intgenes[[r]]<- GenTable(GOdata[[i]], pvalues = resultFis[[i]], orderBy = "pvalues", topNodes=30)
    allRes_intgenes[[r]]$pvalues<-as.numeric(allRes_intgenes[[r]]$pvalues)
    #convert NA values to zero; only if p value is so small it cannot be displayed in r
    allRes_intgenes[[r]][is.na(allRes_intgenes[[r]])]<-0.00000001
    
    #plot GOenrichment 
    david_col=clust.cp.separate
    topGO.outer.layer[[i]]=
      ggplot2::ggplot(allRes_intgenes[[r]], ggplot2::aes(x=(reorder(Term,(-log10(pvalues)))), y=(-log10(pvalues)))) +
      ggplot2::stat_summary(geom = "bar", fun = mean, position = "dodge",
                            col=david_col[i],fill=david_col[i]) +
      ggplot2::coord_flip()+
      ggplot2::xlab(go.type[[r]]) +
      ggplot2::ylab("Enrichment -log10 p-value") +
      ggplot2::labs(title=c("GO enrichment",go.type[[r]]),subtitle=levels(all.markers$cluster)[i])
    
    
    if(r==1)
    {topGO.outer.layer.BP=topGO.outer.layer}else if(r==2)
    {topGO.outer.layer.CC=topGO.outer.layer}else if (r==3)
    {topGO.outer.layer.MF=topGO.outer.layer}  
  }   
} 
save (topGO.outer.layer.BP,topGO.outer.layer.CC,topGO.outer.layer.MF,file='topGO.outer.layer.RData')
}
rm(GOdata.BP.outer.layer,GOdata.CC.outer.layer,GOdata.MF.outer.layer)
```

evaluate the topGO results:
```{r, echo=FALSE, eval=T}
load(file='topGO.outer.layer.RData')
for (i in 1:length(topGO.outer.layer.BP))
  print(topGO.outer.layer.BP[[i]]+
          topGO.outer.layer.MF[[i]]&theme(title = element_text(size=6),axis.text.y = element_text(size=6)))
rm(topGO.outer.layer.MF,topGO.outer.layer.CC,topGO.outer.layer.BP)
```

# inner layer
```{r}
if(cluster.subsets)
{
  data1=data1.subsets$inner.gastrodermis
  data1<-SetIdent(data1,value='IDs.coarse')
  DimPlot(data1, cols = clust.cp.separate,label = T,reduction='umap')+
    NoAxes()
  clusterNames<- read_excel('DataS3.xlsx',
                            sheet = 'InnerClusters') 
    goi = Aa_genes$gene_short_name[match(clusterNames$marker,Aa_genes$name)]
  data1@active.assay='RNA'
  #assign cluster ID to the individual libraries
  data1<-ScaleData(data1,features = goi, split.by = 'orig.ident',verbose=F) #works with scaled gene values, so must scale all of these first (in case they are not within the VarFeatures set!)
  
  clName = vector()
  
  #generate a matrix of values of each cluster for each gene:
  m=AverageExpression(data1,features=goi,slot='scale.data') #Seurat has a function for this
  
  for (j in 1:length(levels(data1@active.ident))) #for each cluster set
  {
    clName[j]=as.integer(which.max(m$RNA[,j])) #choose the highest value
  }
  sort(clName)#for long lists you might want to know what is missing
  clusterNames$ID[clName]
  
  #use the wanted order from the spreadsheet to re-order the clusters:
  #first order the identities..
  data1@active.ident = factor(data1@active.ident,
                              levels(data1@active.ident)[order(clName)])
  #set the names to your IDs; once the gene list works, it is really easy to then update your cluster names with your excel spreadsheet.
  levels(data1@active.ident) = clusterNames$ID[clName][order(clName)]
  #save the IDs in metadata:
  data1@meta.data$IDs = data1@active.ident
  # data1@reductions=data1@reductions[c(1,3,2)]
  print(DimPlot(data1,group.by = c('IDs'),label=F,cols = (clust.cp.separate))+NoAxes())
  
  data1.subsets$inner.gastrodermis=data1
}

```

##### *topGO* | Inner epidermis
```{r, echo=FALSE, eval=F}
if (cluster.subsets)
{
data1=data1.subsets$inner.gastrodermis
all.markers <- FindAllMarkers(data1,
                              return.thresh = 0.01,
                              min.pct = 0.2,
                              only.pos = TRUE,
                              max.cells.per.ident = 100,
                              verbose = F)
library(topGO)

resultFis=NULL
topGO.inner.layer=NULL

Unfold <- Aa_annotations %>% 
  dplyr::mutate(`GOs` = strsplit(as.character(`GOs`), ",")) %>% 
  tidyr::unnest(`GOs`) 
geneID2GO <- Unfold %>% split(x = .$`GOs`, f = .$gene_short_name)
geneNames <- names(geneID2GO)
#set dataset:
all.markers
#process data
GOdata.BP.inner.layer = NULL
GOdata.MF.inner.layer = NULL
GOdata.CC.inner.layer = NULL
myInterestingGenes = NULL
allRes_intgenes = NULL

geneList = NULL
for (r in 1:3)
{
  # this takes a really long time:
  for(i in 1:length(unique(all.markers$cluster)))
  {
    #identify the GOI from the DE list
    myInterestingGenes[[i]] <- all.markers$gene[all.markers$cluster==unique(all.markers$cluster)[i]] #list of genes you want to perform GO enrichment for
    geneList[[i]] <- factor(as.integer(geneNames %in% myInterestingGenes[[i]]))
    names(geneList[[i]]) <- geneNames
    GOdata.BP.inner.layer[[i]] <- new("topGOdata",ontology = "BP", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.MF.inner.layer[[i]] <- new("topGOdata",ontology = "MF", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.CC.inner.layer[[i]] <- new("topGOdata",ontology = "CC", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    go.type=c("Biological Process",'Cellular Compartment','Molecular Function')
    
    if(r==1)
    {GOdata=GOdata.BP.inner.layer}else if(r==2)
    {GOdata=GOdata.CC.inner.layer}else if (r==3)
    {GOdata=GOdata.MF.inner.layer}  
    
    
    #run the test...
    resultFis[[i]] <- runTest(GOdata[[i]], algorithm = "classic", statistic = "fisher")
    pvalFis <- score(resultFis[[i]])
    #filter for only >0.05
    pvalFis = pvalFis[pvalFis>=0.05]
    allRes_intgenes[[r]]<- GenTable(GOdata[[i]], pvalues = resultFis[[i]], orderBy = "pvalues", topNodes=30)
    allRes_intgenes[[r]]$pvalues<-as.numeric(allRes_intgenes[[r]]$pvalues)
    #convert NA values to zero; only if p value is so small it cannot be displayed in r
    allRes_intgenes[[r]][is.na(allRes_intgenes[[r]])]<-0.00000001
    
    #plot GOenrichment 
    david_col=clust.cp.separate
    topGO.inner.layer[[i]]=
      ggplot2::ggplot(allRes_intgenes[[r]], ggplot2::aes(x=(reorder(Term,(-log10(pvalues)))), y=(-log10(pvalues)))) +
      ggplot2::stat_summary(geom = "bar", fun = mean, position = "dodge",
                            col=david_col[i],fill=david_col[i]) +
      ggplot2::coord_flip()+
      ggplot2::xlab(go.type[[r]]) +
      ggplot2::ylab("Enrichment -log10 p-value") +
      ggplot2::labs(title=c("GO enrichment",go.type[[r]]),subtitle=levels(all.markers$cluster)[i])
    
    
    if(r==1)
    {topGO.inner.layer.BP=topGO.inner.layer}else if(r==2)
    {topGO.inner.layer.CC=topGO.inner.layer}else if (r==3)
    {topGO.inner.layer.MF=topGO.inner.layer}  
  }   
} 
save (topGO.inner.layer.BP,topGO.inner.layer.CC,topGO.inner.layer.MF,file='topGO.inner.layer.RData')
rm(GOdata.BP.inner.layer, GO.data,
GOdata.MF.inner.layer,
GOdata.CC.inner.layer)
}
```

evaluate the topGO results:
```{r, echo=FALSE, eval=T}
load(file='topGO.inner.layer.RData')
for (i in 1:length(topGO.inner.layer.BP))
  print(topGO.inner.layer.BP[[i]]+
          topGO.inner.layer.MF[[i]]&theme(title = element_text(size=6),axis.text.y = element_text(size=6)))
rm(topGO.inner.layer.MF,topGO.inner.layer.BP,topGO.inner.layer.CC)
```


# Center
```{r, warning=FALSE}
if(cluster.subsets)
{
  data1=data1.subsets$unchar.center
  data1@active.assay='integrated'
    data1<-SetIdent(data1,value='IDs.coarse')
  DimPlot(data1, cols = clust.cp.separate,label = T,reduction='umap')+
    NoAxes()
  clusterNames<- read_excel('DataS3.xlsx',
                            sheet = 'Center') 
    goi = Aa_genes$gene_short_name[match(clusterNames$marker,Aa_genes$name)]
  data1@active.assay='RNA'
  #assign cluster ID to the individual libraries
  data1<-ScaleData(data1,features = goi, split.by = 'orig.ident',verbose=F) #works with scaled gene values, so must scale all of these first (in case they are not within the VarFeatures set!)
  
  clName = vector()
  
  #generate a matrix of values of each cluster for each gene:
  m=AverageExpression(data1,features=goi,slot='scale.data') #Seurat has a function for this
  
  for (j in 1:length(levels(data1@active.ident))) #for each cluster set
  {
    clName[j]=as.integer(which.max(m$RNA[,j])) #choose the highest value
  }
  sort(clName)#for long lists you might want to know what is missing
  clusterNames$ID[clName]
  
  #use the wanted order from the spreadsheet to re-order the clusters:
  #first order the identities..
  data1@active.ident = factor(data1@active.ident,
                              levels(data1@active.ident)[order(clName)])
  #set the names to your IDs; once the gene list works, it is really easy to then update your cluster names with your excel spreadsheet.
  levels(data1@active.ident) = clusterNames$ID[clName][order(clName)]
  #save the IDs in metadata:
  data1@meta.data$IDs = data1@active.ident
  # data1@reductions=data1@reductions[c(1,3,2)]
  
    data1.subsets$unchar.center=data1
  print(DimPlot(data1,group.by = c('IDs'),label=F,cols = (clust.cp.separate))+NoAxes())
}

```
##Fig.8b.split
```{r}
  data1=data1.subsets$unchar.center
 p1= FeaturePlot(data1,'cytoTRACE.subset', order=T, reduction = 'umap', cols = alpha(rev(magma(50)),0.5),split.by = 'lifehistory')&labs(subtitle = "Differentiation Potential")&NoAxes()+NoLegend()
    p2=FeaturePlot(data1,'cytoTRACE.subset', order=T, reduction = 'umap', cols = alpha(rev(magma(50)),0.5))&labs(tag='B',subtitle = "Differentiation Potential")&NoAxes()+NoLegend()
    p3=FeaturePlot(Aa.Alldata,'cytoTRACE.alldata', order=T, reduction = 'umap.int', cols = alpha(rev(magma(50)),0.5))&labs(tag='A',subtitle = "Differentiation Potential")&NoAxes()+NoLegend()

    pdf('Fig8.pdf',width=12,height = 8)
    p3+p2+p1+plot_layout(design=c(area(1,1),area(1,2),area(2,1,2,2)))
```
##Fig.8C,D
```{r,fig.height=6,fig.width=6}
piwi=genes$gene_short_name[grep('PIWI',genes$Preferred_name)]
nanos=genes$gene_short_name[grep('NANOS',genes$Preferred_name)]
marker.goi=c(
             'PCNA-like-1','HMGB3-like-1','S17AB-like-1','NANO1-like-2',
             'EVE-like-1','AaTWIST-like1','HMA2-like-9','HXA9-like-1',
             'THAP9-like-24','FOXO-like-2','FOXO-like-1',"PIWL2-like-2",
#'KLF6-like-1',
'KLF15-like-1','ARNT-like-1','PITX3-like-1','RX3-like-1',
'NGN2-like-1','ASCL4-like-1','PRD13-like-1','AaMusclebHLH')
 marker.goi=genes$gene_short_name[match(marker.goi,genes$name)]
data1.subsets$unchar.center@active.assay='alra'
Aa.Alldata@active.assay='alra'
# FeaturePlot(data1.subsets$unchar.center,marker.goi[10:12],split.by='lifehistory',order=T,cols=gene.cp)&NoAxes()&NoLegend()
FeaturePlot(Aa.Alldata,marker.goi,order=T,cols=gene.cp)&NoAxes()&NoLegend()
FeaturePlot(data1.subsets$unchar.center,marker.goi,order=T,cols=gene.cp)&NoAxes()&NoLegend()
```
```{r}
pdf('Fig8.CD.pdf',width=8,height = 10,onefile = T)
FeaturePlot(Aa.Alldata,marker.goi,order=T,cols=gene.cp)&NoAxes()&NoLegend()
FeaturePlot(data1.subsets$unchar.center,marker.goi,order=T,cols=gene.cp)&NoAxes()&NoLegend()
dev.off()
```

##### *topGO | center epidermis
```{r, echo=FALSE, eval=F}
if (cluster.subsets)
{
data1=data1.subsets$unchar.center
all.markers <- FindAllMarkers(data1,
                              return.thresh = 0.01,
                              min.pct = 0.2,
                              only.pos = TRUE,
                              max.cells.per.ident = 100,
                              verbose = F)
library(topGO)

resultFis=NULL
topGO.center.layer=NULL

Unfold <- Aa_annotations %>% 
  dplyr::mutate(`GOs` = strsplit(as.character(`GOs`), ",")) %>% 
  tidyr::unnest(`GOs`) 
geneID2GO <- Unfold %>% split(x = .$`GOs`, f = .$gene_short_name)
geneNames <- names(geneID2GO)
#set dataset:
all.markers
#process data
GOdata.BP.center.layer = NULL
GOdata.MF.center.layer = NULL
GOdata.CC.center.layer = NULL
myInterestingGenes = NULL
allRes_intgenes = NULL

geneList = NULL
for (r in 1:3)
{
  # this takes a really long time:
  for(i in 1:length(unique(all.markers$cluster)))
  {
    #identify the GOI from the DE list
    myInterestingGenes[[i]] <- all.markers$gene[all.markers$cluster==unique(all.markers$cluster)[i]] #list of genes you want to perform GO enrichment for
    geneList[[i]] <- factor(as.integer(geneNames %in% myInterestingGenes[[i]]))
    names(geneList[[i]]) <- geneNames
    GOdata.BP.center.layer[[i]] <- new("topGOdata",ontology = "BP", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.MF.center.layer[[i]] <- new("topGOdata",ontology = "MF", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.CC.center.layer[[i]] <- new("topGOdata",ontology = "CC", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    go.type=c("Biological Process",'Cellular Compartment','Molecular Function')
    
    if(r==1)
    {GOdata=GOdata.BP.center.layer}else if(r==2)
    {GOdata=GOdata.CC.center.layer}else if (r==3)
    {GOdata=GOdata.MF.center.layer}  
    
    
    #run the test...
    resultFis[[i]] <- runTest(GOdata[[i]], algorithm = "classic", statistic = "fisher")
    pvalFis <- score(resultFis[[i]])
    #filter for only >0.05
    pvalFis = pvalFis[pvalFis>=0.05]
    allRes_intgenes[[r]]<- GenTable(GOdata[[i]], pvalues = resultFis[[i]], orderBy = "pvalues", topNodes=30)
    allRes_intgenes[[r]]$pvalues<-as.numeric(allRes_intgenes[[r]]$pvalues)
    #convert NA values to zero; only if p value is so small it cannot be displayed in r
    allRes_intgenes[[r]][is.na(allRes_intgenes[[r]])]<-0.00000001
    
    #plot GOenrichment 
    david_col=clust.cp.separate
    topGO.center.layer[[i]]=
      ggplot2::ggplot(allRes_intgenes[[r]], ggplot2::aes(x=(reorder(Term,(-log10(pvalues)))), y=(-log10(pvalues)))) +
      ggplot2::stat_summary(geom = "bar", fun = mean, position = "dodge",
                            col=david_col[i],fill=david_col[i]) +
      ggplot2::coord_flip()+
      ggplot2::xlab(go.type[[r]]) +
      ggplot2::ylab("Enrichment -log10 p-value") +
      ggplot2::labs(title=c("GO enrichment",go.type[[r]]),subtitle=levels(all.markers$cluster)[i])
    
    
    if(r==1)
    {topGO.center.layer.BP=topGO.center.layer}else if(r==2)
    {topGO.center.layer.CC=topGO.center.layer}else if (r==3)
    {topGO.center.layer.MF=topGO.center.layer}  
  }   
} 
save (topGO.center.layer.BP,topGO.center.layer.CC,topGO.center.layer.MF,file='topGO.center.layer.RData')
rm(GOdata.BP.center.layer,GOdata,
GOdata.MF.center.layer,
GOdata.CC.center.layer,
myInterestingGenes,
allRes_intgenes)
}
```

evaluate the topGO results:
```{r, echo=FALSE, eval=T, fig.width=12}
load(file='topGO.center.layer.RData')
for (i in 1:length(topGO.center.layer.BP))
  print(topGO.center.layer.BP[[i]]+
          topGO.center.layer.MF[[i]]&theme(title = element_text(size=6),axis.text.y = element_text(size=6)))
rm(topGO.center.layer.MF,topGO.center.layer.BP,topGO.center.layer.CC)
```
```{r}
goi = c('AaTWIST-like1','FOXA2-like-1')
piwi=genes$gene_short_name[grep('PIWI',genes$Preferred_name)]
nanos=genes$gene_short_name[grep('NANOS',genes$Preferred_name)]
data1=data1.subsets$unchar.center
  data1@active.assay='alra'
  plot(FeaturePlot(data1,c('FOXA2-like-1','AaBrachyury','AaTWIST-like1',piwi[c(3,4)],nanos[1:2],"S17AB-like-1","ERG-like-2"),order=T,cols=gene.cp,label=F)&NoAxes()&NoLegend())
  Aa.Alldata@active.assay = 'alra'
  plot(  FeaturePlot(Aa.Alldata,c('FOXA2-like-1','AaBrachyury','AaTWIST-like1',piwi[c(3,4)],nanos[1:2],"S17AB-like-1","ERG-like-2"),order=T,cols=gene.cp,label=F)&NoAxes()&NoLegend())
```


# Dig.Gland
```{r,warning=F,eval=FALSE}
if(cluster.subsets)
{
  data1=data1.subsets$gland.dig
  data1<-SetIdent(data1,value = 'IDs.coarse')        
  clusterNames<- read_excel('DataS3.xlsx',
                            sheet = 'Dig.Gland')
    goi = Aa_genes$gene_short_name[match(clusterNames$marker,Aa_genes$name)]
  data1@active.assay='RNA'
  data1<-ScaleData(data1,features = goi,verbose=F)
  clName = vector()
  
  #generate a matrix of values of each cluster for each gene:
  m=AverageExpression(data1,features=goi,slot='scale.data') #Seurat has a function for this
  
  for (j in 1:length(levels(data1@active.ident))) #for each cluster set
  {
    clName[j]=as.integer(which.max(m$RNA[,j])) #choose the highest value
  }
  sort(clName)#for long lists you might want to know what is missing
  clusterNames$ID[clName]
  DimPlot(data1,cols=clust.cp.separate)
  #use the wanted order from the spreadsheet to re-order the clusters:
  #first order the identities..
  data1@active.ident = factor(data1@active.ident,
                              levels(data1@active.ident)[order(clName)])
  #set the names to your IDs; once the gene list works, it is really easy to then update your cluster names with your excel spreadsheet.
  levels(data1@active.ident) = clusterNames$ID[clName][order(clName)]
  #save the IDs in metadata:
  data1@meta.data$IDs = data1@active.ident
  data1@active.assay='integrated'
  # data1@reductions=data1@reductions[c(1,3,2)]
  print(DimPlot(data1,group.by = c('IDs'),label=F,reduction = 'umap',
                cols = (clust.cp.separate))+NoAxes())
  
  data1.subsets$gland.dig=data1
}
```

```{r,eval=F}
proteinases<- read_excel("Aaur2.func.Annotation.xlsx", sheet= 'serine proteases')
proteinases<-Aa_genes$gene_short_name[match(proteinases$names,Aa_genes$name)]
#need to load the markers here for this to work: HAS NOT BEEN CALCULATED YET!
# goi = proteinases$gene_short_names[which(rowSums(data1.subsets$gland.dig@assays$RNA@counts[proteinases$gene_short_names,])>=100)]
goi=intersect(markers$gland.dig$gene,proteinases$gene_short_names)
DotPlot(data1.subsets$gland.dig,'RNA',goi,scale.by='size', 
        # col.min = 0, col.max = 3,  
        cols = c('lightgrey','darkred')) + 
  RotatedAxis() +FontSize(6,6)
```

```{r, fig.height=6}
goi=c('AaASCL5-like','ASCL5-like-3','ASL1B-like-2')
# not expressed: 'ASL1B-like-1','ASCL5-like-2', Neural:'ASCL5-like-2',
goi=Aa_genes$gene_short_name[match(goi,Aa_genes$name)]
# ID.separate HAS NOT BEEN CALCULATED YET
DotPlot(Aa.Alldata,'RNA',goi,group.by = 'ID.separate',scale.by='radius', 
        cols = c('lightgrey','darkred'))&RotatedAxis()
```

```{r}
data1=data1.subsets$gland.dig
data1@active.assay='alra'
FeaturePlot(data1, c('AaASCL5-like','ASCL5-like-3'),order=T,cols=gene.cp,ncol=1)&NoAxes()&NoLegend()
```



##### *topGO | digestive gland cells
```{r, echo=FALSE, eval=F}
if (cluster.subsets)
{
data1=data1.subsets$gland.dig
all.markers <- FindAllMarkers(data1,
                              return.thresh = 0.01,
                              min.pct = 0.2,
                              only.pos = TRUE,
                              max.cells.per.ident = 100,
                              verbose = F)
library(topGO)

resultFis=NULL
topGO.digestive.layer=NULL

Unfold <- Aa_annotations %>% 
  dplyr::mutate(`GOs` = strsplit(as.character(`GOs`), ",")) %>% 
  tidyr::unnest(`GOs`) 
geneID2GO <- Unfold %>% split(x = .$`GOs`, f = .$gene_short_name)
geneNames <- names(geneID2GO)
#set dataset:
all.markers
#process data
GOdata.BP.digestive.layer = NULL
GOdata.MF.digestive.layer = NULL
GOdata.CC.digestive.layer = NULL
myInterestingGenes = NULL
allRes_intgenes = NULL

geneList = NULL
for (r in 1:3)
{
  # this takes a really long time:
  for(i in 1:length(unique(all.markers$cluster)))
  {
    #identify the GOI from the DE list
    myInterestingGenes[[i]] <- all.markers$gene[all.markers$cluster==unique(all.markers$cluster)[i]] #list of genes you want to perform GO enrichment for
    geneList[[i]] <- factor(as.integer(geneNames %in% myInterestingGenes[[i]]))
    names(geneList[[i]]) <- geneNames
    GOdata.BP.digestive.layer[[i]] <- new("topGOdata",ontology = "BP", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.MF.digestive.layer[[i]] <- new("topGOdata",ontology = "MF", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.CC.digestive.layer[[i]] <- new("topGOdata",ontology = "CC", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    go.type=c("Biological Process",'Cellular Compartment','Molecular Function')
    
    if(r==1)
    {GOdata=GOdata.BP.digestive.layer}else if(r==2)
    {GOdata=GOdata.CC.digestive.layer}else if (r==3)
    {GOdata=GOdata.MF.digestive.layer}  
    
    
    #run the test...
    resultFis[[i]] <- runTest(GOdata[[i]], algorithm = "classic", statistic = "fisher")
    pvalFis <- score(resultFis[[i]])
    #filter for only >0.05
    pvalFis = pvalFis[pvalFis>=0.05]
    allRes_intgenes[[r]]<- GenTable(GOdata[[i]], pvalues = resultFis[[i]], orderBy = "pvalues", topNodes=30)
    allRes_intgenes[[r]]$pvalues<-as.numeric(allRes_intgenes[[r]]$pvalues)
    #convert NA values to zero; only if p value is so small it cannot be displayed in r
    allRes_intgenes[[r]][is.na(allRes_intgenes[[r]])]<-0.00000001
    
    #plot GOenrichment 
    david_col=clust.cp.separate
    topGO.digestive.layer[[i]]=
      ggplot2::ggplot(allRes_intgenes[[r]], ggplot2::aes(x=(reorder(Term,(-log10(pvalues)))), y=(-log10(pvalues)))) +
      ggplot2::stat_summary(geom = "bar", fun = mean, position = "dodge",
                            col=david_col[i],fill=david_col[i]) +
      ggplot2::coord_flip()+
      ggplot2::xlab(go.type[[r]]) +
      ggplot2::ylab("Enrichment -log10 p-value") +
      ggplot2::labs(title=c("GO enrichment",go.type[[r]]),subtitle=levels(all.markers$cluster)[i])
    
    
    if(r==1)
    {topGO.digestive.layer.BP=topGO.digestive.layer}else if(r==2)
    {topGO.digestive.layer.CC=topGO.digestive.layer}else if (r==3)
    {topGO.digestive.layer.MF=topGO.digestive.layer}  
  }   
} 
save (topGO.digestive.layer.BP,topGO.digestive.layer.CC,topGO.digestive.layer.MF,file='topGO.digestive.layer.RData')
rm(GOdata.BP.digestive.layer,
GOdata.MF.digestive.layer,
GOdata.CC.digestive.layer,
myInterestingGenes,
allRes_intgenes)
}
```

evaluate the topGO results:
```{r, echo=FALSE, eval=T}
load(file='topGO.digestive.layer.RData')
for (i in 1:length(topGO.digestive.layer.BP))
  print(topGO.digestive.layer.BP[[i]]+
          topGO.digestive.layer.MF[[i]]&theme(title = element_text(size=6),axis.text.y = element_text(size=6)))
rm(topGO.digestive.layer.MF,topGO.digestive.layer.BP,topGO.digestive.layer.CC)
```

## Neural 1
```{r }
if(cluster.subsets)
{
  data1=data1.subsets$neural.1
  # data1@reductions = data1@reductions[c(1,4,3,2)]
 
  clusterNames<- read_excel('DataS3.xlsx', sheet = 'Neuro')
                            # sheet = 'Neuro')
  
    goi = Aa_genes$gene_short_name[match(clusterNames$marker,Aa_genes$name)]
    data1<-SetIdent(data1,value = 'IDs.fine')
   print(DimPlot(data1,cols=clust.cp.separate,reduction='umap'))
  data1@active.assay='RNA'

   #assign cluster ID to the individual libraries
  data1<-ScaleData(data1,features = goi,verbose=F)
  clName = vector()
  
  #generate a matrix of values of each cluster for each gene:
  m=AverageExpression(data1,features=goi,slot='scale.data') #Seurat has a function for this
  
  for (j in 1:length(levels(data1@active.ident))) #for each cluster set
  {
    clName[j]=as.integer(which.max(m$RNA[,j])) #choose the highest value
  }
  sort(clName)#for long lists you might want to know what is missing
  clusterNames$ID[clName]
  
  #use the wanted order from the spreadsheet to re-order the clusters:
  #first order the identities..
  data1@active.ident = factor(data1@active.ident,
                              levels(data1@active.ident)[order(clName)])
  #set the names to your IDs; once the gene list works, it is really easy to then update your cluster names with your excel spreadsheet.
  levels(data1@active.ident) = clusterNames$ID[clName][order(clName)]
  #save the IDs in metadata:
  data1@meta.data$IDs = data1@active.ident
  data1@active.assay='integrated'
  # data1@reductions=data1@reductions[c(1,3,2)]
  data1.subsets$neural.1=data1
  
  print(DimPlot(data1,group.by = c('IDs'),label=F,reduction = 'umap',
          cols = (clust.cp.separate))+NoAxes()+labs(title = 'neural.1 population'))
}
```

```{r}
data1=data1.subsets$neural.1
  ids.cluster.library = as.data.frame(table(Idents(data1), data1@meta.data$orig.ident))
  colnames(ids.cluster.library) = c('ID','Library','CellCount')
  clust.cp =(clust.cp.separate)
  print(
    DimPlot(data1, label = F,label.size = 5, repel = T,
            cols = clust.cp)+NoLegend()+NoAxes()+
      ggplot(ids.cluster.library, aes(fill=ID, y= log(CellCount),
                                      x=Library)) +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),
                            x=(Library)),
               position="stack", stat="identity", width = 0.5)+
      scale_fill_manual(values = clust.cp)+
      theme(axis.text.x = element_text(#face="bold", color="#993333", 
        size=8, angle=-45,hjust=0,vjust = 0.5))+
      geom_area(mapping =aes(fill=ID, y= log(CellCount),
                             x=as.integer(Library)),
                position="stack", stat="identity",alpha=0.2 , size=.5, colour="white") +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),#this re-plots the bars over the area
                            x=(Library)),
               position="stack", stat="identity", width = 0.5)+labs(title = 'Cell distribution',subtitle = 'neurons')
  )
```


##### topGO | Neural Clusters
```{r, echo=FALSE, eval=F,warning=FALSE}
if(cluster.subsets)
{
data1=data1.subsets$neural.1
all.markers <- FindAllMarkers(data1,
                              return.thresh = 0.01,
                              min.pct = 0.2,
                              only.pos = TRUE,
                              max.cells.per.ident = 100,
                              verbose = F)
library(topGO)

resultFis=NULL
topGO.neurons=NULL

Unfold <- Aa_annotations %>% 
  dplyr::mutate(`GOs` = strsplit(as.character(`GOs`), ",")) %>% 
  tidyr::unnest(`GOs`) 
geneID2GO <- Unfold %>% split(x = .$`GOs`, f = .$gene_short_name)
geneNames <- names(geneID2GO)
#set dataset:
all.markers
#process data
GOdata.BP.neurons = NULL
GOdata.MF.neurons = NULL
GOdata.CC.neurons = NULL
myInterestingGenes = NULL
allRes_intgenes = NULL

geneList = NULL
for (r in 1:3)
{
  # this takes a really long time:
  for(i in 1:length(unique(all.markers$cluster)))
  {
    #identify the GOI from the DE list
    myInterestingGenes[[i]] <- all.markers$gene[all.markers$cluster==unique(all.markers$cluster)[i]] #list of genes you want to perform GO enrichment for
    geneList[[i]] <- factor(as.integer(geneNames %in% myInterestingGenes[[i]]))
    names(geneList[[i]]) <- geneNames
    GOdata.BP.neurons[[i]] <- new("topGOdata",ontology = "BP", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.MF.neurons[[i]] <- new("topGOdata",ontology = "MF", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.CC.neurons[[i]] <- new("topGOdata",ontology = "CC", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    go.type=c("Biological Process",'Cellular Compartment','Molecular Function')
    
    if(r==1)
    {GOdata=GOdata.BP.neurons}else if(r==2)
    {GOdata=GOdata.CC.neurons}else if (r==3)
    {GOdata=GOdata.MF.neurons}  
    
    
    #run the test...
    resultFis[[i]] <- runTest(GOdata[[i]], algorithm = "classic", statistic = "fisher")
    pvalFis <- score(resultFis[[i]])
    #filter for only >0.05
    pvalFis = pvalFis[pvalFis>=0.05]
    allRes_intgenes[[r]]<- GenTable(GOdata[[i]], pvalues = resultFis[[i]], orderBy = "pvalues", topNodes=30)
    allRes_intgenes[[r]]$pvalues<-as.numeric(allRes_intgenes[[r]]$pvalues)
    #convert NA values to zero; only if p value is so small it cannot be displayed in r
    allRes_intgenes[[r]][is.na(allRes_intgenes[[r]])]<-0.00000001
    
    #plot GOenrichment 
    david_col=clust.cp.separate
    topGO.neurons[[i]]=
      ggplot2::ggplot(allRes_intgenes[[r]], ggplot2::aes(x=(reorder(Term,(-log10(pvalues)))), y=(-log10(pvalues)))) +
      ggplot2::stat_summary(geom = "bar", fun = mean, position = "dodge",
                            col=david_col[i],fill=david_col[i]) +
      ggplot2::coord_flip()+
      ggplot2::xlab(go.type[[r]]) +
      ggplot2::ylab("Enrichment -log10 p-value") +
      ggplot2::labs(title=c("GO enrichment",go.type[[r]]),subtitle=levels(all.markers$cluster)[i])
    
    
    if(r==1)
    {topGO.neurons.BP=topGO.neurons}else if(r==2)
    {topGO.neurons.CC=topGO.neurons}else if (r==3)
    {topGO.neurons.MF=topGO.neurons}  
  }   
} 
save (topGO.neurons.BP,topGO.neurons.CC,topGO.neurons.MF,file='topGO.neurons.RData')
}
```

evaluate the topGO results:
```{r, echo=FALSE, eval=T}
load(file='topGO.neurons.RData')
for (i in 1:length(topGO.neurons.BP))
  print(topGO.neurons.BP[[i]]+
          topGO.neurons.MF[[i]]&theme(title = element_text(size=6),axis.text.y = element_text(size=6)))
rm(topGO.neurons.BP,topGO.neurons.CC,topGO.neurons.MF)
```

```{r}
data1.subsets$all.neurons$nClass = 'NA'
nC1=WhichCells(data1.subsets$all.neurons,idents = levels(data1.subsets$all.neurons)[10:18])
nC2=WhichCells(data1.subsets$all.neurons,idents = levels(data1.subsets$all.neurons)[1:9])
data1.subsets$all.neurons$nClass[nC1] = 'n1'
data1.subsets$all.neurons$nClass[nC2] = 'n2'
data1=SetIdent(data1.subsets$all.neurons,value='nClass')
neur.markers <- FindAllMarkers(data1,
                                  logfc.threshold = 1,
                                 # features = intersect(AaTF_list,rownames(data1)),
                                 min.pct = 0.8,
                                 only.pos = TRUE,
                                 max.cells.per.ident = 200,
                                 return.thresh = 0.001,verbose=F)
```

```{r}
data1=data1.subsets$all.neurons
data1@active.ident=factor(data1@active.ident,levels(data1@active.ident)[rev(1:19)])
 print(DotPlot(data1,'RNA',c('ins1a-like-1'),cols=c('lightgrey','#8C6D14'),scale.by = 'size',dot.min = 0.01,col.min = 0,scale = F)&RotatedAxis())#
  print(DotPlot(data1,'RNA',c('ashA'),cols=c('lightgrey','purple4'),scale.by = 'size',dot.min = 0.01,col.min = 0,scale = F)&RotatedAxis())#
```

#### Neural 2
```{r }
if(cluster.subsets)
{
  data1=data1.subsets$neural.2
  # data1@reductions = data1@reductions[c(1,4,3,2)]
  print(DimPlot(data1,cols=clust.cp.separate,reduction='umap'))
  clusterNames<- read_excel('DataS3.xlsx',
                            sheet = 'Neuro')
  
    goi = Aa_genes$gene_short_name[match(clusterNames$marker,Aa_genes$name)]
  data1@active.assay='RNA'
  data1<-SetIdent(data1,value = 'IDs.fine')
  #assign cluster ID to the individual libraries
  data1<-ScaleData(data1,features = goi,verbose=F)
  clName = vector()
  
  #generate a matrix of values of each cluster for each gene:
  m=AverageExpression(data1,features=goi,slot='scale.data') #Seurat has a function for this
  
  for (j in 1:length(levels(data1@active.ident))) #for each cluster set
  {
    clName[j]=as.integer(which.max(m$RNA[,j])) #choose the highest value
  }
  sort(clName)#for long lists you might want to know what is missing
  clusterNames$ID[clName]
  
  #use the wanted order from the spreadsheet to re-order the clusters:
  #first order the identities..
  data1@active.ident = factor(data1@active.ident,
                              levels(data1@active.ident)[order(clName)])
  #set the names to your IDs; once the gene list works, it is really easy to then update your cluster names with your excel spreadsheet.
  levels(data1@active.ident) = clusterNames$ID[clName][order(clName)]
  #save the IDs in metadata:
  data1@meta.data$IDs = data1@active.ident
  data1@active.assay='integrated'
  # data1@reductions=data1@reductions[c(1,3,2)]
  
  data1.subsets$neural.2=data1
  #rev(clust.cp.graded)[2:24]
  print(DimPlot(data1,group.by = c('IDs'),label=F,reduction = 'umap',
          cols = clust.cp.separate)+NoAxes()+labs(title = 'neural population'))

}
```

```{r}
data1=data1.subsets$neural.2
  ids.cluster.library = as.data.frame(table(Idents(data1), data1@meta.data$orig.ident))
  colnames(ids.cluster.library) = c('ID','Library','CellCount')
  clust.cp = rev(clust.cp.graded)
  print(
    DimPlot(data1, label = F,label.size = 5, repel = T,
            cols = clust.cp)+NoLegend()+NoAxes()+
      ggplot(ids.cluster.library, aes(fill=ID, y= log(CellCount),
                                      x=Library)) +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),
                            x=(Library)),
               position="stack", stat="identity", width = 0.5)+
      scale_fill_manual(values = clust.cp)+
      theme(axis.text.x = element_text(#face="bold", color="#993333", 
        size=8, angle=-45,hjust=0,vjust = 0.5))+
      geom_area(mapping =aes(fill=ID, y= log(CellCount),
                             x=as.integer(Library)),
                position="stack", stat="identity",alpha=0.2 , size=.5, colour="white") +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),#this re-plots the bars over the area
                            x=(Library)),
               position="stack", stat="identity", width = 0.5)+labs(title = 'Cell distribution',subtitle = 'neurons')
  )
```


##### topGO | Neural Clusters
```{r, echo=FALSE, eval=F,warning=FALSE,message=FALSE}
if(cluster.subsets)
{
data1=data1.subsets$neural.2
all.markers <- FindAllMarkers(data1,
                              return.thresh = 0.01,
                              min.pct = 0.2,
                              only.pos = TRUE,
                              max.cells.per.ident = 100,
                              verbose = F)
library(topGO)

resultFis=NULL
topGO.neuron2=NULL

Unfold <- Aa_annotations %>% 
  dplyr::mutate(`GOs` = strsplit(as.character(`GOs`), ",")) %>% 
  tidyr::unnest(`GOs`) 
geneID2GO <- Unfold %>% split(x = .$`GOs`, f = .$gene_short_name)
geneNames <- names(geneID2GO)
#set dataset:
all.markers
#process data
GOdata.BP.neuron2 = NULL
GOdata.MF.neuron2 = NULL
GOdata.CC.neuron2 = NULL
myInterestingGenes = NULL
allRes_intgenes = NULL

geneList = NULL
for (r in 1:3)
{
  # this takes a really long time:
  for(i in 1:length(unique(all.markers$cluster)))
  {
    #identify the GOI from the DE list
    myInterestingGenes[[i]] <- all.markers$gene[all.markers$cluster==unique(all.markers$cluster)[i]] #list of genes you want to perform GO enrichment for
    geneList[[i]] <- factor(as.integer(geneNames %in% myInterestingGenes[[i]]))
    names(geneList[[i]]) <- geneNames
    GOdata.BP.neuron2[[i]] <- new("topGOdata",ontology = "BP", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.MF.neuron2[[i]] <- new("topGOdata",ontology = "MF", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.CC.neuron2[[i]] <- new("topGOdata",ontology = "CC", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    go.type=c("Biological Process",'Cellular Compartment','Molecular Function')
    
    if(r==1)
    {GOdata=GOdata.BP.neuron2}else if(r==2)
    {GOdata=GOdata.CC.neuron2}else if (r==3)
    {GOdata=GOdata.MF.neuron2}  
    
    
    #run the test...
    resultFis[[i]] <- runTest(GOdata[[i]], algorithm = "classic", statistic = "fisher")
    pvalFis <- score(resultFis[[i]])
    #filter for only >0.05
    pvalFis = pvalFis[pvalFis>=0.05]
    allRes_intgenes[[r]]<- GenTable(GOdata[[i]], pvalues = resultFis[[i]], orderBy = "pvalues", topNodes=30)
    allRes_intgenes[[r]]$pvalues<-as.numeric(allRes_intgenes[[r]]$pvalues)
    #convert NA values to zero; only if p value is so small it cannot be displayed in r
    allRes_intgenes[[r]][is.na(allRes_intgenes[[r]])]<-0.00000001
    
    #plot GOenrichment 
    david_col=clust.cp.separate
    topGO.neurons[[i]]=
      ggplot2::ggplot(allRes_intgenes[[r]], ggplot2::aes(x=(reorder(Term,(-log10(pvalues)))), y=(-log10(pvalues)))) +
      ggplot2::stat_summary(geom = "bar", fun = mean, position = "dodge",
                            col=david_col[i],fill=david_col[i]) +
      ggplot2::coord_flip()+
      ggplot2::xlab(go.type[[r]]) +
      ggplot2::ylab("Enrichment -log10 p-value") +
      ggplot2::labs(title=c("GO enrichment",go.type[[r]]),subtitle=levels(all.markers$cluster)[i])
    
    
    if(r==1)
    {topGO.neuron2.BP=topGO.neurons}else if(r==2)
    {topGO.neuron2.CC=topGO.neurons}else if (r==3)
    {topGO.neuron2.MF=topGO.neurons}  
  }   
} 
save (topGO.neuron2.BP,topGO.neuron2.CC,topGO.neuron2.MF,file='topGO.neuron2.RData')
}
```

evaluate the topGO results:
```{r, echo=FALSE, eval=T}
load(file='topGO.neuron2.RData')
for (i in 1:length(topGO.neuron2.BP))
  print(topGO.neuron2.BP[[i]]+
          topGO.neuron2.MF[[i]]&theme(title = element_text(size=6),axis.text.y = element_text(size=6)))
rm(topGO.neuron2.MF,topGO.neuron2.BP,topGO.neuron2.CC)
```

# Neurons combined
```{r}
  data1=subset(Aa.Alldata,idents=c('neural.1','neural.2'))
    data1@active.assay = 'integrated'
    
    data1 <- FindVariableFeatures(data1,nfeatures = 2000,verbose = F)#
    #what if you use the approach from the nematostella analysis, where you import the scaled data from the full dataset
    # data1 <- ScaleData(data1,verbose = F)#,split.by = 'orig.ident')
    # #use the full dataset scaling:
    Aa.Alldata@active.assay='integrated'
    coi = colnames(data1)
    t=ScaleData(Aa.Alldata,features = data1@assays$RNA@var.features,verbose=F)
    data1@assays$integrated@scale.data = t@assays$integrated@scale.data[,coi]
    rm(t)
    data1 <- RunPCA(data1, pcs.compute = 50,verbose = F)
    ElbowPlot(object = data1, ndims = 50)
    # Determine percent of variation associated with each PC
    pct <- data1[["pca"]]@stdev / sum(data1[["pca"]]@stdev) * 100
    # Calculate cumulative percents for each PC
    cumu <- cumsum(pct)
    # Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
    co1 <- which(cumu > 90 & pct < 5)[1]
    # Determine the difference between variation of PC and subsequent PC: last point where change of % of variation is more than 0.1%
    co2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1
    # choose the minimum of these two metrics as the PCs covering the majority of the variation in the data
    pcs <- min(co1, co2)
    d = 1:20
    data1 <- RunUMAP(data1, reduction ="pca", n.neighbors = 10L,spread =1,seed.use = 0, 
                     dims = d,reduction.name ='umap',
                     reduction.key ='umap',min.dist = 0.6,
                     local.connectivity = 10,verbose = F)#
    data1 <- FindNeighbors(object = data1,reduction ="pca",dims = d,
                           nn.method = 'annoy',  annoy.metric = 'cosine',verbose = F)
    data1 <- FindClusters(object = data1,resolution = 2,random.seed = 0,verbose = F)
    # #look at relationship between clusters
    data1 <- BuildClusterTree(object = data1, reorder = TRUE,
                              reorder.numeric = TRUE,verbose = F)#, dims = c(1:d))
    data1$IDs.fine = data1@active.ident
    data1 <- FindClusters(object = data1,resolution = 0.2,random.seed = 0,verbose = F)
    # #look at relationship between clusters
    data1 <- BuildClusterTree(object = data1, reorder = TRUE,
                              reorder.numeric = TRUE,verbose = F)#, dims = c(1:d))
    data1$IDs.coarse = data1@active.ident
    data1@reductions=data1@reductions[c(1,3,2)]
    data1.subsets$all.neurons = data1
```
```{r}
data1.subsets$all.neurons$orig.ident =as.factor(data1.subsets$all.neurons$orig.ident) 

data1.subsets$all.neurons$orig.ident=factor(data1.subsets$all.neurons$orig.ident,levels(data1.subsets$all.neurons$orig.ident)[c(7,9,8,10,1:6)])
```


labels the clusters
```{r }
if(cluster.subsets)
{
data1=data1.subsets$all.neurons
  clusterNames<- read_excel('DataS3.xlsx',
                            sheet = 'Neuro')
  
    goi = Aa_genes$gene_short_name[match(clusterNames$marker,Aa_genes$name)]
    data1<-SetIdent(data1,value = 'IDs.fine')
   print(DimPlot(data1,cols=clust.cp.separate,reduction='umap'))
  data1@active.assay='RNA'

   #assign cluster ID to the individual libraries
  data1<-ScaleData(data1,features = goi,verbose=F)
  clName = vector()
  
  #generate a matrix of values of each cluster for each gene:
  m=AverageExpression(data1,features=goi,slot='scale.data') #Seurat has a function for this
  
  for (j in 1:length(levels(data1@active.ident))) #for each cluster set
  {
    clName[j]=as.integer(which.max(m$RNA[,j])) #choose the highest value
  }
  sort(clName)#for long lists you might want to know what is missing
  clusterNames$ID[clName]
  
  #use the wanted order from the spreadsheet to re-order the clusters:
  #first order the identities..
  data1@active.ident = factor(data1@active.ident,
                              levels(data1@active.ident)[order(clName)])
  #set the names to your IDs; once the gene list works, it is really easy to then update your cluster names with your excel spreadsheet.
  levels(data1@active.ident) = clusterNames$ID[clName][order(clName)]
  #save the IDs in metadata:
  data1@meta.data$IDs = data1@active.ident
  data1@active.assay='integrated'
  # data1@reductions=data1@reductions[c(1,3,2)]
  data1.subsets$all.neurons=data1
  
  print(DimPlot(data1,group.by = c('IDs'),label=F,reduction = 'umap',
          cols = (clust.cp.separate))+NoAxes()+labs(title = 'all neural population'))
}
```

```{r, fig.height=6}
data1=data1.subsets$all.neurons
  ids.cluster.library = as.data.frame(table(Idents(data1), data1@meta.data$orig.ident))
  colnames(ids.cluster.library) = c('ID','Library','CellCount')
  clust.cp =(clust.cp.separate)
  print(
    DimPlot(data1, label = F,label.size = 5, repel = T,
            cols = clust.cp)+NoLegend()+NoAxes()+
      ggplot(ids.cluster.library, aes(fill=ID, y= log(CellCount),
                                      x=Library)) +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),
                            x=(Library)),
               position="stack", stat="identity", width = 0.5)+
      scale_fill_manual(values = clust.cp)+
      theme(axis.text.x = element_text(#face="bold", color="#993333", 
        size=8, angle=-45,hjust=0,vjust = 0.5))+
      geom_area(mapping =aes(fill=ID, y= log(CellCount),
                             x=as.integer(Library)),
                position="stack", stat="identity",alpha=0.2 , size=.5, colour="white") +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),#this re-plots the bars over the area
                            x=(Library)),
               position="stack", stat="identity", width = 0.5)+labs(title = 'Cell distribution',subtitle = 'neurons')
  )
```

```{r}
data1.subsets$all.neurons$nClass = 'NA'
nC1=colnames(data1.subsets$neural.1)
nC2=colnames(data1.subsets$neural.2)
data1.subsets$all.neurons$nClass[nC1] = 'n1'
data1.subsets$all.neurons$nClass[nC2] = 'n2'
data1=SetIdent(data1.subsets$all.neurons,value='nClass')
neur.markers <- FindAllMarkers(data1,
                                  logfc.threshold = 1,
                                 # features = intersect(AaTF_list,rownames(data1)),
                                 min.pct = 0.8,
                                 only.pos = TRUE,
                                 max.cells.per.ident = 200,
                                 return.thresh = 0.001,verbose=F)
```

```{r, fig.height=8}
data1.subsets$neural.1$all.n.IDs = 'NA'
data1.subsets$neural.2$all.n.IDs = 'NA'
nCa=colnames(data1.subsets$all.neurons)
nC1=colnames(data1.subsets$neural.1)
nC2=colnames(data1.subsets$neural.2)
data1.subsets$neural.1$all.n.IDs[nC1] = as.character(data1.subsets$all.neurons@active.ident[nC1])
data1.subsets$neural.2$all.n.IDs[nC2] = as.character(data1.subsets$all.neurons@active.ident[nC2])

data1.subsets$neural.1$all.n.IDs = as.factor(data1.subsets$neural.1$all.n.IDs)
data1.subsets$neural.2$all.n.IDs =as.factor(data1.subsets$neural.2$all.n.IDs)

DimPlot(data1.subsets$neural.1,cols=neur.cp,group.by = 'all.n.IDs')+NoAxes()+labs(title='n1 Class neurons')&theme(legend.position = 'bottom')&
  DimPlot(data1.subsets$neural.2,cols=neur.cp,group.by = 'all.n.IDs')+NoAxes()+labs(title='n2 Class neurons')&theme(legend.position = 'bottom')
```

## plot neurons
```{r, fig.height=12}
# data1=SetIdent(Aa.Alldata,value='ID.separate')
DimPlot(data1.subsets$neural.1,cols=c(stepped(8)[5:8],stepped3(8)[5:8],'orange','darkorange2'))+NoAxes()+labs(title='n1 Class neurons')&theme(legend.position = 'bottom')&
  DimPlot(data1.subsets$neural.2,cols=c("#33a02c",stepped2(8)[5:8],stepped(12)[10:12],'darkgreen'))+NoAxes()+labs(title='n2 Class neurons')&theme(legend.position = 'bottom')
```
## Fig4a.subset
```{r, fig.width=4,fig.height=4}
# data1=SetIdent(Aa.Alldata,value='ID.separate')
names(neur.cp)=levels(data1.subsets$all.neurons)

pdf('Fig4ai.pdf',width = 6,height = 6)
DimPlot(data1.subsets$all.neurons,cols=neur.cp,reduction='umap')+NoAxes()+labs(title='All neurons')+NoLegend()
dev.off()

DimPlot(data1.subsets$all.neurons,cols=neur.cp,reduction='umap')+NoAxes()+labs(title='All neurons')+NoLegend()
```
```{r, fig.height=8}
# add the distriubtion plot
data1=data1.subsets$all.neurons
 ids.cluster.library = as.data.frame(table(Idents(data1), data1@meta.data$orig.ident))
  colnames(ids.cluster.library) = c('ID','Library','CellCount')
 ids.cluster.library$ID<-factor(ids.cluster.library$ID,levels(ids.cluster.library$ID)[rev(1:length(levels(data1.subsets$all.neurons)))])

   clust.cp=rev(neur.cp)
  print(
      ggplot(ids.cluster.library, aes(fill=ID, y= log(CellCount),
                                      x=Library)) +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),
                            x=(Library)),
               position="stack", stat="identity", width = 0.5)+
      scale_fill_manual(values = clust.cp)+
      theme(axis.text.x = element_text(#face="bold", color="#993333", 
        size=8, angle=-45,hjust=0,vjust = 0.5))+
      geom_area(mapping =aes(fill=ID, y= log(CellCount),
                             x=as.integer(Library)),
                position="stack", stat="identity",alpha=0.2 , size=.5, colour="white") +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),#this re-plots the bars over the area
                            x=(Library)),
               position="stack", stat="identity", width = 0.5)+labs(title = 'neurons')
  )
```
## Fig4B
```{r, fig.height=6,fig.width=6}
# add the distriubtion plot
data1=data1.subsets$all.neurons
 ids.cluster.library = as.data.frame(table(Idents(data1), data1@meta.data$orig.ident))
  colnames(ids.cluster.library) = c('ID','Library','CellCount')
 ids.cluster.library$ID<-factor(ids.cluster.library$ID,levels(ids.cluster.library$ID)[rev(1:length(levels(data1.subsets$all.neurons)))])

   clust.cp=rev(neur.cp)
  print(
      ggplot(ids.cluster.library, aes(fill=ID, y= log(CellCount),
                                      x=Library)) +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),
                            x=(Library)),
               position="fill", stat="identity", width = 0.5)+
      scale_fill_manual(values = clust.cp)+
      theme(axis.text.x = element_text(#face="bold", color="#993333", 
        size=10, angle=-45,hjust=0,vjust = 0.5))+
      geom_area(mapping =aes(fill=ID, y= log(CellCount),
                             x=as.integer(Library)),
                position="fill", stat="identity",alpha=0.2 , size=.5, colour="white") +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),#this re-plots the bars over the area
                            x=(Library)),
               position="fill", stat="identity", width = 0.5)+labs(title = 'neurons')
  )
```

```{r}
pdf('Fig4B.pdf',width = 6,height = 6)
      ggplot(ids.cluster.library, aes(fill=ID, y= log(CellCount),
                                      x=Library)) +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),
                            x=(Library)),
               position="fill", stat="identity", width = 0.5)+
      scale_fill_manual(values = clust.cp)+
      theme(axis.text.x = element_text(#face="bold", color="#993333", 
        size=10, angle=-45,hjust=0,vjust = 0.5))+
      geom_area(mapping =aes(fill=ID, y= log(CellCount),
                             x=as.integer(Library)),
                position="fill", stat="identity",alpha=0.2 , size=.5, colour="white") +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),#this re-plots the bars over the area
                            x=(Library)),
               position="fill", stat="identity", width = 0.5)+labs(title = 'neurons')
dev.off()
```

## Fig4A.fulldata
``` {r,fig.height=4,fig.width=4}
coi=NULL
for(c in 1:length(levels(data1.subsets$all.neurons)))
  coi[[c]]=WhichCells(data1.subsets$all.neurons,idents=levels(data1.subsets$all.neurons)[c])
names(coi)=levels(data1.subsets$all.neurons)
DimPlot(Aa.Alldata,cells.highlight = coi,cols.highlight = rev(neur.cp[sort(names(coi))]))&NoAxes()&NoLegend()
```

#Fig4C
```{r}
# data1=SetIdent(Aa.Alldata,value='ID.separate')[47:64]
data1=data1.subsets$all.neurons
data1@active.assay
# data1@active.ident=factor(data1@active.ident,levels(data1@active.ident)[rev(1:14)])
 print(DotPlot(data1,'RNA',c('ins1a-like-1'),cols=c('lightgrey','darkorange3'),scale.by = 'size',dot.min = 0.01,col.min = 0,scale = F, idents =levels(data1))&RotatedAxis())#
  print(DotPlot(data1,'RNA',c('ashA'),idents =levels(data1),cols=c('lightgrey','darkgreen'),scale.by = 'size',dot.min = 0.01,col.min = 0,scale = F)&RotatedAxis())#
```

```{r}
pdf('Fig4C.pdf',height = 10,width = 10)
 DotPlot(data1,'RNA',c('ins1a-like-1'),cols=c('lightgrey','darkorange3'),scale.by = 'size',dot.min = 0.01,col.min = 0,scale = F, idents =levels(data1))&RotatedAxis()&
DotPlot(data1,'RNA',c('ashA'),idents =levels(data1),cols=c('lightgrey','darkgreen'),scale.by = 'size',dot.min = 0.01,col.min = 0,scale = F)&RotatedAxis()
 dev.off()
```


```{r}
 print(DotPlot(Aa.Alldata,'RNA',
               # 'ARNT-like-1',
               # goi,
               c('pou4-like1','tba1-like3','tbb-like-1'),#,'CPT1A-like-1','AaChitinase1','CPT1A-like-2','CPT1A-like-3'),
               group.by = 'ID.separate',cols=c('lightgrey','darkred'),col.min = 0,scale.by = 'radius')&RotatedAxis()&coord_flip())
#,'AaTBA1-like3','KCNA2-like-10','TBB-like-1','TBB-like-2'
```

Neurons only:
```{r,warning=FALSE,fig.height=6}
s=10
   list = NULL
    for (c in  1:length(levels(data1.subsets[[s]]@active.ident)))
    {
      x=markers[[s]][as.numeric(markers[[s]]$cluster)==c,][1:min(5,length(which(as.numeric(markers[[s]]$cluster)==c))),7]
      list=c(list,x)
    }
    
goi1=all.markers.Alldata$gene[which(all.markers.Alldata$cluster == 'neural.1')][3:11]
goi2=all.markers.Alldata$gene[which(all.markers.Alldata$cluster == 'neural.2')][1:5]
DotPlot(data1.subsets[[s]],'RNA', unique(c(goi2,list,goi1)),scale.by='radius' ,dot.min = 0.1, col.min = 0,cols = c('lightgrey','darkred'))+RotatedAxis() +FontSize(6,8)+ labs(title=names(data1.subsets)[s],subtitle = 'DEGs p.val <=0.001')+theme(axis.text.y = element_text(colour=all.clusters.cp[levels(data1.subsets[[s]])],size=8))
DotPlot(SetIdent(Aa.Alldata,value='ID.separate'),'RNA', unique(c('ASCL4-like-1',goi2,'AaPOU4-like1',list,goi1)),scale.by='radius' ,dot.min = 0.1, col.min = 0,cols = c('lightgrey','darkred'))+RotatedAxis() +FontSize(6,8)+ labs(title=names(data1.subsets)[s],subtitle = 'DEGs p.val <=0.001')+theme(axis.text.y = element_text(colour=all.clusters.cp,size=8),legend.position = 'bottom')
```

#Fig.4D
Neurons only:
```{r,warning=FALSE,fig.height=8,fig.width=12}

if(!exists('all.markers.Alldata'))
load(file = 'AllData.integrated2.markers.RObj')
goi1=all.markers.Alldata$gene[which(all.markers.Alldata$cluster == 'neural.1')][3:15]
goi2=all.markers.Alldata$gene[which(all.markers.Alldata$cluster == 'neural.2')][1:5]

s=10
   list = NULL
    for (c in  1:length(levels(data1.subsets[[s]]@active.ident)))
    {
            x=markers[[s]][as.numeric(markers[[s]]$cluster)==c,][1:min(5,length(which(as.numeric(markers[[s]]$cluster)==c))),7]

      y=markers.TF[[s]][as.numeric(markers.TF[[s]]$cluster)==c,][1:min(5,length(which(as.numeric(markers.TF[[s]]$cluster)==c))),7]
      list=c(list,x,y)
    }

DotPlot(data1.subsets[[s]],'RNA', unique(c(goi2,'tbb-like-1',list,goi1)),scale.by='radius' ,dot.min = 0.1, col.min = 0,cols = c('lightgrey','darkred'))+RotatedAxis() +FontSize(6,10)+ labs(title=names(data1.subsets)[s],subtitle = 'DEGs p.val <=0.001')+theme(axis.text.y = element_text(colour=neur.cp,size=8),legend.position = 'bottom',legend.text = (element_text(size=8)))
```
```{r}
pdf('fig4D.pdf',width=15,height = 5)
DotPlot(data1.subsets[[s]],'RNA', unique(c(goi2,'tbb-like-1',list,goi1)),scale.by='radius' ,dot.min = 0.1, col.min = 0,cols = c('lightgrey','darkred'))+RotatedAxis() +FontSize(6,10)+ labs(title=names(data1.subsets)[s],subtitle = 'DEGs p.val <=0.001')+theme(axis.text.y = element_text(colour=neur.cp,size=8),legend.position = 'bottom',legend.text = (element_text(size=8)))
dev.off()
```

##### topGO | Neural Clusters
```{r, echo=FALSE, eval=F,warning=FALSE,message=FALSE}
if(cluster.subsets)
{
data1=data1.subsets$all.neurons
all.markers <- FindAllMarkers(data1,
                              return.thresh = 0.01,
                              min.pct = 0.2,
                              only.pos = TRUE,
                              max.cells.per.ident = 100,
                              verbose = F)
library(topGO)

resultFis=NULL
topGO.neurons=NULL

Unfold <- Aa_annotations %>% 
  dplyr::mutate(`GOs` = strsplit(as.character(`GOs`), ",")) %>% 
  tidyr::unnest(`GOs`) 
geneID2GO <- Unfold %>% split(x = .$`GOs`, f = .$gene_short_name)
geneNames <- names(geneID2GO)
#set dataset:
all.markers
#process data
GOdata.BP.all.neurons = NULL
GOdata.MF.all.neurons = NULL
GOdata.CC.all.neurons = NULL
myInterestingGenes = NULL
allRes_intgenes = NULL

geneList = NULL
for (r in 1:3)
{
  # this takes a really long time:
  for(i in 1:length(unique(all.markers$cluster)))
  {
    #identify the GOI from the DE list
    myInterestingGenes[[i]] <- all.markers$gene[all.markers$cluster==unique(all.markers$cluster)[i]] #list of genes you want to perform GO enrichment for
    geneList[[i]] <- factor(as.integer(geneNames %in% myInterestingGenes[[i]]))
    names(geneList[[i]]) <- geneNames
    GOdata.BP.all.neurons[[i]] <- new("topGOdata",ontology = "BP", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.MF.all.neurons[[i]] <- new("topGOdata",ontology = "MF", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.CC.all.neurons[[i]] <- new("topGOdata",ontology = "CC", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    go.type=c("Biological Process",'Cellular Compartment','Molecular Function')
    
    if(r==1)
    {GOdata=GOdata.BP.all.neurons}else if(r==2)
    {GOdata=GOdata.CC.all.neurons}else if (r==3)
    {GOdata=GOdata.MF.all.neurons}  
    
    
    #run the test...
    resultFis[[i]] <- runTest(GOdata[[i]], algorithm = "classic", statistic = "fisher")
    pvalFis <- score(resultFis[[i]])
    #filter for only >0.05
    pvalFis = pvalFis[pvalFis>=0.05]
    allRes_intgenes[[r]]<- GenTable(GOdata[[i]], pvalues = resultFis[[i]], orderBy = "pvalues", topNodes=30)
    allRes_intgenes[[r]]$pvalues<-as.numeric(allRes_intgenes[[r]]$pvalues)
    #convert NA values to zero; only if p value is so small it cannot be displayed in r
    allRes_intgenes[[r]][is.na(allRes_intgenes[[r]])]<-0.00000001
    
    #plot GOenrichment 
    david_col=clust.cp.separate
    topGO.neurons[[i]]=
      ggplot2::ggplot(allRes_intgenes[[r]], ggplot2::aes(x=(reorder(Term,(-log10(pvalues)))), y=(-log10(pvalues)))) +
      ggplot2::stat_summary(geom = "bar", fun = mean, position = "dodge",
                            col=david_col[i],fill=david_col[i]) +
      ggplot2::coord_flip()+
      ggplot2::xlab(go.type[[r]]) +
      ggplot2::ylab("Enrichment -log10 p-value") +
      ggplot2::labs(title=c("GO enrichment",go.type[[r]]),subtitle=levels(all.markers$cluster)[i])
    
    
    if(r==1)
    {topGO.all.neurons.BP=topGO.neurons}else if(r==2)
    {topGO.all.neurons.CC=topGO.neurons}else if (r==3)
    {topGO.all.neurons.MF=topGO.neurons}  
  }   
} 
save (topGO.all.neurons.BP,topGO.all.neurons.CC,topGO.all.neurons.MF,file='topGO.all.neurons.RData')
rm(GOdata.BP.all.neurons,GOdata,
GOdata.MF.all.neurons ,GOdata.CC.all.neurons,
myInterestingGenes,
allRes_intgenes)
}
```

evaluate the topGO results:
```{r, echo=FALSE, eval=T}
load(file='topGO.all.neurons.RData')
for (i in 1:length(topGO.all.neurons.BP))
  print(topGO.all.neurons.BP[[i]]+
          topGO.all.neurons.MF[[i]]&theme(title = element_text(size=6),axis.text.y = element_text(size=6)))
rm(topGO.all.neurons.MF,topGO.all.neurons.BP,topGO.all.neurons.CC)
```

names require review:
# Mucin
```{r, warning=FALSE}
if(cluster.subsets)
{
  data1=data1.subsets$gland.muc
  data1<-SetIdent(data1,value = 'IDs.coarse')#fine picks up small clusters but don't know what to do with these
  print(DimPlot(data1,cols = clust.cp.separate,label = T)&NoAxes())
  # data1@reductions <- data1@reductions[c(1,3,2)]
  clusterNames<- read_excel('DataS3.xlsx',
                            sheet = 'MucinClusters')
  
    goi = Aa_genes$gene_short_name[match(clusterNames$marker,Aa_genes$name)]
  data1@active.assay='RNA'
  #assign cluster ID to the individual libraries
  data1<-ScaleData(data1,features = goi, split.by = 'orig.ident',verbose=F) #works with scaled gene values, so must scale all of these first (in case they are not within the VarFeatures set!)
  
  clName = vector()
  
  #generate a matrix of values of each cluster for each gene:
  m=AverageExpression(data1,features=goi,slot='scale.data') #Seurat has a function for this
  
  for (j in 1:length(levels(data1@active.ident))) #for each cluster set
  {
    clName[j]=as.integer(which.max(m$RNA[,j])) #choose the highest value
  }
  sort(clName) #this prints which IDs were selected; sanity check that it did what you wanted... should have no unwanted replicated IDs
  clusterNames$ID[clName] #you can also look at the names to see where the problem might have been
  
  DimPlot(data1,label = T,cols=clust.cp.separate,reduction='umap')+NoAxes()
  
  #use the wanted order from the spreadsheet to re-order the clusters:
  #first order the identities..
  data1@active.ident = factor(data1@active.ident,
                              levels(data1@active.ident)[order(clName)])
  #set the names to your IDs; once the gene list works, it is really easy to then update your cluster names with your excel spreadsheet.
  levels(data1@active.ident) = clusterNames$ID[clName][order(clName)]
  #save the IDs in metadata:
  data1@meta.data$IDs = data1@active.ident
  data1@active.assay='integrated'
    data1.subsets$gland.muc =data1

  print(DimPlot(data1,group.by = c('IDs'),label=F,cols = (clust.cp.separate))+NoAxes())
  }
```

##### topGO | mucin Clusters
```{r, echo=FALSE, eval=F,message=FALSE}
if(cluster.subsets)
{data1=data1.subsets$gland.muc
all.markers <- FindAllMarkers(data1,
                              return.thresh = 0.01,
                              min.pct = 0.2,
                              only.pos = TRUE,
                              max.cells.per.ident = 100,
                              verbose = F)
library(topGO)

resultFis=NULL
topGO.mucin=NULL

Unfold <- Aa_annotations %>% 
  dplyr::mutate(`GOs` = strsplit(as.character(`GOs`), ",")) %>% 
  tidyr::unnest(`GOs`) 
geneID2GO <- Unfold %>% split(x = .$`GOs`, f = .$gene_short_name)
geneNames <- names(geneID2GO)
#set dataset:
all.markers
#process data
GOdata.BP.mucin = NULL
GOdata.MF.mucin = NULL
GOdata.CC.mucin = NULL
myInterestingGenes = NULL
allRes_intgenes = NULL

geneList = NULL
for (r in 1:3)
{
  # this takes a really long time:
  for(i in 1:length(unique(all.markers$cluster)))
  {
    #identify the GOI from the DE list
    myInterestingGenes[[i]] <- all.markers$gene[all.markers$cluster==unique(all.markers$cluster)[i]] #list of genes you want to perform GO enrichment for
    geneList[[i]] <- factor(as.integer(geneNames %in% myInterestingGenes[[i]]))
    names(geneList[[i]]) <- geneNames
    GOdata.BP.mucin[[i]] <- new("topGOdata",ontology = "BP", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.MF.mucin[[i]] <- new("topGOdata",ontology = "MF", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.CC.mucin[[i]] <- new("topGOdata",ontology = "CC", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    go.type=c("Biological Process",'Cellular Compartment','Molecular Function')
    
    if(r==1)
    {GOdata=GOdata.BP.mucin}else if(r==2)
    {GOdata=GOdata.CC.mucin}else if (r==3)
    {GOdata=GOdata.MF.mucin}  
    
    
    #run the test...
    resultFis[[i]] <- runTest(GOdata[[i]], algorithm = "classic", statistic = "fisher")
    pvalFis <- score(resultFis[[i]])
    #filter for only >0.05
    pvalFis = pvalFis[pvalFis>=0.05]
    allRes_intgenes[[r]]<- GenTable(GOdata[[i]], pvalues = resultFis[[i]], orderBy = "pvalues", topNodes=30)
    allRes_intgenes[[r]]$pvalues<-as.numeric(allRes_intgenes[[r]]$pvalues)
    #convert NA values to zero; only if p value is so small it cannot be displayed in r
    allRes_intgenes[[r]][is.na(allRes_intgenes[[r]])]<-0.00000001
    
    #plot GOenrichment 
    david_col=clust.cp.separate
    topGO.mucin[[i]]=
      ggplot2::ggplot(allRes_intgenes[[r]], ggplot2::aes(x=(reorder(Term,(-log10(pvalues)))), y=(-log10(pvalues)))) +
      ggplot2::stat_summary(geom = "bar", fun = mean, position = "dodge",
                            col=david_col[i],fill=david_col[i]) +
      ggplot2::coord_flip()+
      ggplot2::xlab(go.type[[r]]) +
      ggplot2::ylab("Enrichment -log10 p-value") +
      ggplot2::labs(title=c("GO enrichment",go.type[[r]]),subtitle=levels(all.markers$cluster)[i])
    
    
    if(r==1)
    {topGO.mucin.BP=topGO.mucin}else if(r==2)
    {topGO.mucin.CC=topGO.mucin}else if (r==3)
    {topGO.mucin.MF=topGO.mucin}  
  }   
} 
save (topGO.mucin.BP,topGO.mucin.CC,topGO.mucin.MF,file='topGO.mucin.RData')
}
```

evaluate the topGO results:
```{r, echo=FALSE, eval=T}
load(file='topGO.mucin.RData')
for (i in 1:length(topGO.mucin.BP))
  print(topGO.mucin.BP[[i]]+
          topGO.mucin.MF[[i]]&theme(title = element_text(size=6),axis.text.y = element_text(size=6)))
rm(topGO.mucin.MF,topGO.mucin.BP,topGO.mucin.CC)
```

# Cnidocytes All
```{r }
if(cluster.subsets)
{

  data1=data1.subsets$cnidocyte #
  data1<-SetIdent(data1,value = 'IDs.coarse')
  
  clusterNames<- read_excel('DataS3.xlsx',
                            sheet = 'CnidoClusters')
  
    goi = Aa_genes$gene_short_name[match(clusterNames$marker,Aa_genes$name)]
  data1@active.assay='RNA'
  #assign cluster ID to the individual libraries

  data1<-ScaleData(data1,features = goi, split.by='orig.ident',verbose=F) #works with scaled gene values, so must scale all of these first (in case they are not within the VarFeatures set!)
  clName = vector()
  
  #generate a matrix of values of each cluster for each gene:
  m=AverageExpression(data1,features=goi,slot='scale.data',verbose=F) #Seurat has a function for this
  
  for (j in 1:length(levels(data1@active.ident))) #for each cluster set
  {
    clName[j]=as.integer(which.max(m$RNA[,j])) #choose the highest value
  }
  
  sort(clName) #this prints which IDs were selected; sanity check that it did what you wanted... should have no unwanted replicated IDs
  clusterNames$ID[clName] #you can also look at the names to see where the problem might have been
  plot(DimPlot(data1,label = T,cols=clust.cp.separate,reduction='umap')+NoAxes())
  
  #use the wanted order from the spreadsheet to re-order the clusters:
  #first order the identities..
  data1@active.ident = factor(data1@active.ident,
                              levels(data1@active.ident)[order(clName)])
  #set the names to your IDs; once the gene list works, it is really easy to then update your cluster names with your excel spreadsheet.
  levels(data1@active.ident) = clusterNames$ID[clName][order(clName)]
  #save the IDs in metadata:
  data1@meta.data$IDs = data1@active.ident
  
  # cni.goi.nv=c('SOX4-like-1','FOS-like-4','GFI1B-like-1','MYC-like-5',
  # 'ZN431-like-1','AaJUN-like3','MYC-like-4','SX18A-like-1')
  # print(FeaturePlot(data1,cni.goi.nv,order = T,cols = gene.cp.dark,label = F,ncol=4,reduction='umap')&
  #   NoAxes()&NoLegend())
  
  data1@active.assay='integrated'
  data1.subsets$cnidocyte=data1
  print(DimPlot(data1,cols=clust.cp.separate,reduction='umap',group.by = 'IDs')+NoAxes())#+
          # DimPlot(data1,cols=clust.cp.separate,reduction='umap',group.by = 'IDs.alldata')&NoAxes())
}
```
##### *topGO | Cnidocyte
```{r, echo=FALSE, eval=F}
if (cluster.subsets)
{
data1=data1.subsets$cnidocyte
all.markers <- FindAllMarkers(data1,
                              return.thresh = 0.01,
                              min.pct = 0.2,
                              only.pos = TRUE,
                              max.cells.per.ident = 100,
                              verbose = F)
library(topGO)

resultFis=NULL
topGO.cnidocyte=NULL

Unfold <- Aa_annotations %>% 
  dplyr::mutate(`GOs` = strsplit(as.character(`GOs`), ",")) %>% 
  tidyr::unnest(`GOs`) 
geneID2GO <- Unfold %>% split(x = .$`GOs`, f = .$gene_short_name)
geneNames <- names(geneID2GO)
#set dataset:
all.markers
#process data
GOdata.BP.cnidocyte = NULL
GOdata.MF.cnidocyte = NULL
GOdata.CC.cnidocyte = NULL
myInterestingGenes = NULL
allRes_intgenes = NULL

geneList = NULL
for (r in 1:3)
{
  # this takes a really long time:
  for(i in 1:length(unique(all.markers$cluster)))
  {
    #identify the GOI from the DE list
    myInterestingGenes[[i]] <- all.markers$gene[all.markers$cluster==unique(all.markers$cluster)[i]] #list of genes you want to perform GO enrichment for
    geneList[[i]] <- factor(as.integer(geneNames %in% myInterestingGenes[[i]]))
    names(geneList[[i]]) <- geneNames
    GOdata.BP.cnidocyte[[i]] <- new("topGOdata",ontology = "BP", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.MF.cnidocyte[[i]] <- new("topGOdata",ontology = "MF", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.CC.cnidocyte[[i]] <- new("topGOdata",ontology = "CC", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    go.type=c("Biological Process",'Cellular Compartment','Molecular Function')
    
    if(r==1)
    {GOdata=GOdata.BP.cnidocyte}else if(r==2)
    {GOdata=GOdata.CC.cnidocyte}else if (r==3)
    {GOdata=GOdata.MF.cnidocyte}  
    
    
    #run the test...
    resultFis[[i]] <- runTest(GOdata[[i]], algorithm = "classic", statistic = "fisher")
    pvalFis <- score(resultFis[[i]])
    #filter for only >0.05
    pvalFis = pvalFis[pvalFis>=0.05]
    allRes_intgenes[[r]]<- GenTable(GOdata[[i]], pvalues = resultFis[[i]], orderBy = "pvalues", topNodes=30)
    allRes_intgenes[[r]]$pvalues<-as.numeric(allRes_intgenes[[r]]$pvalues)
    #convert NA values to zero; only if p value is so small it cannot be displayed in r
    allRes_intgenes[[r]][is.na(allRes_intgenes[[r]])]<-0.00000001
    
    #plot GOenrichment 
    david_col=clust.cp.separate
    topGO.cnidocyte[[i]]=
      ggplot2::ggplot(allRes_intgenes[[r]], ggplot2::aes(x=(reorder(Term,(-log10(pvalues)))), y=(-log10(pvalues)))) +
      ggplot2::stat_summary(geom = "bar", fun = mean, position = "dodge",
                            col=david_col[i],fill=david_col[i]) +
      ggplot2::coord_flip()+
      ggplot2::xlab(go.type[[r]]) +
      ggplot2::ylab("Enrichment -log10 p-value") +
      ggplot2::labs(title=c("GO enrichment",go.type[[r]]),subtitle=levels(all.markers$cluster)[i])
    
    
    if(r==1)
    {topGO.cnidocyte.BP=topGO.cnidocyte}else if(r==2)
    {topGO.cnidocyte.CC=topGO.cnidocyte}else if (r==3)
    {topGO.cnidocyte.MF=topGO.cnidocyte}  
  }   
} 
save (topGO.cnidocyte.BP,topGO.cnidocyte.CC,topGO.cnidocyte.MF,file='topGO.cnidocyte.RData')

rm(GOdata.BP.cnidocyte,GOdata,
GOdata.MF.cnidocyte ,GOdata.CC.cnidocyte,
myInterestingGenes,
allRes_intgenes, topGO.cnidocyte)
}
```

evaluate the topGO results:
```{r, echo=FALSE, eval=T}
load(file='topGO.cnidocyte.RData')
for (i in 1:length(topGO.cnidocyte.BP))
  print(topGO.cnidocyte.BP[[i]]+
          topGO.cnidocyte.MF[[i]]&theme(title = element_text(size=6),axis.text.y = element_text(size=6)))
rm(topGO.cnidocyte.BP,topGO.cnidocyte.MF,topGO.cnidocyte.CC)
```

# Cnidocyte specification
```{r}
goi=c("SOX4-like-1","MYC-like-7",  "PRD13-like-1", "ZN431-like-1", 
 "COT2-like-1","FOSX-like-1", "AaJUN-like-2", 'FOS-like-4','AL-like-1',
"AaJUN-like3", "ZN333-like-2",  
"GFI1B-like-1", "FOXN4-like-1", "MYC-like-5","AaPOU4-like1", 
 'MYC-like-4',"BGB-like-1") 
goi=Aa_genes$gene_short_name[match(goi,Aa_genes$name)]

data1=data1.subsets$cnidocyte

data1@active.assay='alra'
FeaturePlot(data1,goi,order=T,cols=gene.cp)&NoAxes()&NoLegend()

```
## Fig3c
```{r}
toxins=c('NAS4-like-10','NAS6-like-8','NAS4-like-11','AaNAS4-like','NAS15-like-18','GGH-like-8','NAS15-like-3','NAS4-like-21','NAS6-like-6')
toxins=Aa_genes$gene_short_name[match(toxins,Aa_genes$name)]
  print(
        DotPlot(data1,'RNA', features = toxins,
                              scale.by='radius' , col.min = 0, col.max = 4,  
                              cols = c('lightgrey','red'))+
      RotatedAxis() +FontSize(6,8) +labs(titel='DETFs'))
  # data1@active.assay='RNA'
  data1.subsets$cnidocyte@active.assay='alra'
  pdf('Fig3c.pdf',width = 12,height = 8)
FeaturePlot(data1.subsets$cnidocyte,toxins[c(5:7,1:3)],order=T,cols=c('lightgrey','red'),ncol=3)&NoAxes()&NoLegend()
 dev.off()
```

##Fig3B
```{r, fig.height=8,fig.width=6,warning=F}
goi=c("SOX4-like-1","MYC-like-7",  "PRD13-like-1", "ZN431-like-1", 
 "COT2-like-1","FOSX-like-1", "AaJUN-like-2", 'FOS-like-4','AL-like-1',
"AaJUN-like3", "ZN333-like-2",  
"GFI1B-like-1", "FOXN4-like-1", "MYC-like-5","AaPOU4-like1", 
 'MYC-like-4',"BGB-like-1") 
goi=Aa_genes$gene_short_name[match(goi,Aa_genes$name)]
pdf('Fig.3B.pdf',height = 8,width = 6)
DotPlot(Aa.Alldata,'RNA',goi,group.by = 'ID.separate',cols = c('lightgrey','red'),col.min = 0)+FontSize(8,10) +theme(axis.text.y = element_text(colour = all.clusters.cp),legend.title = element_text(size = 6),legend.text = element_text(size=6))+RotatedAxis()
dev.off()
DotPlot(Aa.Alldata,'RNA',goi,group.by = 'ID.separate',cols = c('lightgrey','red'),col.min = 0)+FontSize(8,10) +theme(axis.text.y = element_text(colour = all.clusters.cp),legend.title = element_text(size = 6),legend.text = element_text(size=6))+RotatedAxis()

```
##fig3A
``` {r, fig.width=6,warning=F}

clust.cp=all.clusters.cp[levels(data1.subsets$cnidocyte)]
clust.cp[3]='green4'
clust.cp[4]='violet'
clust.cp[8]='pink4'

cnido.clust=DimPlot(data1.subsets$cnidocyte,cols=clust.cp)+NoAxes()+NoLegend()
  data1=data1.subsets$cnidocyte
  ids.cluster.library = as.data.frame(table(Idents(data1), data1@meta.data$lifehistory.tissue))
  colnames(ids.cluster.library) = c('ID','Stage','CellCount')
  
  print(cnido.clust+
      ggplot(ids.cluster.library, aes(fill=ID, y= log(CellCount),
                                      x=Stage)) +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),
                            x=(Stage)),
               position="fill", stat="identity", width = 0.5)+
      scale_fill_manual(values = clust.cp)+
      theme(axis.text.x = element_text(#face="bold", color="#993333", 
        size=8, angle=-45,hjust=0,vjust = 0.5))+
      geom_area(mapping =aes(fill=ID, y= log(CellCount),
                             x=as.integer(Stage)),
                position="fill", stat="identity",alpha=0.2 , size=.5, colour="white") +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),#this re-plots the bars over the area
                            x=(Stage)),
               position="fill", stat="identity", width = 0.5)
  )

```
```{r}
pdf('Fig3A.pdf',height = 6,width = 8)
print(cnido.clust+
      ggplot(ids.cluster.library, aes(fill=ID, y= log(CellCount),
                                      x=Stage)) +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),
                            x=(Stage)),
               position="fill", stat="identity", width = 0.5)+
      scale_fill_manual(values = clust.cp)+
      theme(axis.text.x = element_text(#face="bold", color="#993333", 
        size=8, angle=-45,hjust=0,vjust = 0.5))+
      geom_area(mapping =aes(fill=ID, y= log(CellCount),
                             x=as.integer(Stage)),
                position="fill", stat="identity",alpha=0.2 , size=.5, colour="white") +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),#this re-plots the bars over the area
                            x=(Stage)),
               position="fill", stat="identity", width = 0.5)
  )
dev.off()
```

##fig3A.inset

``` {r,fig.height=4,fig.width=4}
coi=NULL
for(c in 1:length(levels(data1.subsets$cnidocyte)))
  coi[[c]]=WhichCells(data1.subsets$cnidocyte,idents=levels(data1.subsets$cnidocyte)[c])
names(coi)=levels(data1.subsets$cnidocyte)
pdf('Fig3Ai.pdf',height = 5,width = 5)
DimPlot(Aa.Alldata,cells.highlight = coi,cols.highlight = rev(clust.cp[sort(names(coi))]))&NoAxes()&NoLegend()
dev.off()
```
# Muscle.st : OLIVER
Output of the standard things as with other clusters, but Oliver is looking into the details of this cluster.

```{r }
if(cluster.subsets)
{
  data1=data1.subsets[[1]]
  data1<-SetIdent(data1,value = 'IDs.coarse')
  # data1@reductions = data1@reductions[c(1,4,3,2)]
  print(DimPlot(data1,cols=clust.cp.separate,reduction='umap'))
  
  clusterNames<- read_excel('DataS3.xlsx',
                            sheet = 'Muscle.ST') 
  
    goi = Aa_genes$gene_short_name[match(clusterNames$marker,Aa_genes$name)]
  data1@active.assay='RNA'
  
  #assign cluster ID to the individual libraries
  data1<-ScaleData(data1,features = goi,verbose=F)#, split.by = 'orig.ident') #works with scaled gene values, so must scale all of these first (in case they are not within the VarFeatures set!)
  
  clName = vector()
  
  #generate a matrix of values of each cluster for each gene:
  m=AverageExpression(data1,features=goi,slot='scale.data') #Seurat has a function for this
  
  for (j in 1:length(levels(data1@active.ident))) #for each cluster set
  {
    clName[j]=as.integer(which.max(m$RNA[,j])) #choose the highest value
  }
  sort(clName)#for long lists you might want to know what is missing
  clusterNames$ID[clName]
  
  #use the wanted order from the spreadsheet to re-order the clusters:
  #first order the identities..
  data1@active.ident = factor(data1@active.ident,
                              levels(data1@active.ident)[order(clName)])
  #set the names to your IDs; once the gene list works, it is really easy to then update your cluster names with your excel spreadsheet.
  levels(data1@active.ident) = clusterNames$ID[clName][order(clName)]
  #save the IDs in metadata:
  data1@meta.data$IDs = data1@active.ident
  data1@active.assay='integrated'
  # data1@reductions=data1@reductions[c(1,3,2)]
  print(DimPlot(data1,group.by = c('IDs'),label=F,reduction = 'umap',
                cols = (clust.cp.separate))+NoAxes())
  
  data1.subsets[[1]]=data1
}
```
##### *topGO | st.muscle
```{r, echo=FALSE, eval=F}
if (cluster.subsets)
{
data1=data1.subsets$muscle.st
all.markers <- FindAllMarkers(data1,
                              return.thresh = 0.01,
                              min.pct = 0.2,
                              only.pos = TRUE,
                              max.cells.per.ident = 100,
                              verbose = F)
library(topGO)

resultFis=NULL
topGO.st.muscle=NULL

Unfold <- Aa_annotations %>% 
  dplyr::mutate(`GOs` = strsplit(as.character(`GOs`), ",")) %>% 
  tidyr::unnest(`GOs`) 
geneID2GO <- Unfold %>% split(x = .$`GOs`, f = .$gene_short_name)
geneNames <- names(geneID2GO)
#set dataset:
all.markers
#process data
GOdata.BP.st.muscle = NULL
GOdata.MF.st.muscle = NULL
GOdata.CC.st.muscle = NULL
myInterestingGenes = NULL
allRes_intgenes = NULL

geneList = NULL
for (r in 1:3)
{
  # this takes a really long time:
  for(i in 1:length(unique(all.markers$cluster)))
  {
    #identify the GOI from the DE list
    myInterestingGenes[[i]] <- all.markers$gene[all.markers$cluster==unique(all.markers$cluster)[i]] #list of genes you want to perform GO enrichment for
    geneList[[i]] <- factor(as.integer(geneNames %in% myInterestingGenes[[i]]))
    names(geneList[[i]]) <- geneNames
    GOdata.BP.st.muscle[[i]] <- new("topGOdata",ontology = "BP", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.MF.st.muscle[[i]] <- new("topGOdata",ontology = "MF", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    GOdata.CC.st.muscle[[i]] <- new("topGOdata",ontology = "CC", allGenes = geneList[[i]],annot = annFUN.gene2GO, gene2GO = geneID2GO)
    go.type=c("Biological Process",'Cellular Compartment','Molecular Function')
    
    if(r==1)
    {GOdata=GOdata.BP.st.muscle}else if(r==2)
    {GOdata=GOdata.CC.st.muscle}else if (r==3)
    {GOdata=GOdata.MF.st.muscle}  
    
    
    #run the test...
    resultFis[[i]] <- runTest(GOdata[[i]], algorithm = "classic", statistic = "fisher")
    pvalFis <- score(resultFis[[i]])
    #filter for only >0.05
    pvalFis = pvalFis[pvalFis>=0.05]
    allRes_intgenes[[r]]<- GenTable(GOdata[[i]], pvalues = resultFis[[i]], orderBy = "pvalues", topNodes=30)
    allRes_intgenes[[r]]$pvalues<-as.numeric(allRes_intgenes[[r]]$pvalues)
    #convert NA values to zero; only if p value is so small it cannot be displayed in r
    allRes_intgenes[[r]][is.na(allRes_intgenes[[r]])]<-0.00000001
    
    #plot GOenrichment 
    david_col=clust.cp.separate
    topGO.st.muscle[[i]]=
      ggplot2::ggplot(allRes_intgenes[[r]], ggplot2::aes(x=(reorder(Term,(-log10(pvalues)))), y=(-log10(pvalues)))) +
      ggplot2::stat_summary(geom = "bar", fun = mean, position = "dodge",
                            col=david_col[i],fill=david_col[i]) +
      ggplot2::coord_flip()+
      ggplot2::xlab(go.type[[r]]) +
      ggplot2::ylab("Enrichment -log10 p-value") +
      ggplot2::labs(title=c("GO enrichment",go.type[[r]]),subtitle=levels(all.markers$cluster)[i])
    
    
    if(r==1)
    {topGO.st.muscle.BP=topGO.st.muscle}else if(r==2)
    {topGO.st.muscle.CC=topGO.st.muscle}else if (r==3)
    {topGO.st.muscle.MF=topGO.st.muscle}  
  }   
} 
save (topGO.st.muscle.BP,topGO.st.muscle.CC,topGO.st.muscle.MF,file='topGO.st.muscle.RData')

rm(GOdata.BP.st.muscle,GOdata,
GOdata.MF.st.muscle ,GOdata.CC.st.muscle,
myInterestingGenes,
allRes_intgenes)
}
```

evaluate the topGO results:
```{r, echo=FALSE, eval=T}
load(file='topGO.st.muscle.RData')
for (i in 1:length(topGO.st.muscle.BP))
  print(topGO.st.muscle.BP[[i]]+
          topGO.st.muscle.MF[[i]]&theme(title = element_text(size=6),axis.text.y = element_text(size=6)))
rm(topGO.st.muscle.MF,topGO.st.muscle.BP,topGO.st.muscle.CC)
```

ran the cytotrace on all possibilities separately. Use this as importation for Monocle: ran also random UMAP seed 10 times via cluster.
#Monocle3

#AllData Monocle ----
```{r}
library(monocle3,quietly = T)
data1=Aa.Alldata
data1@active.assay='integrated'
data1@reductions$umap=data1@reductions$umap.int

cds <- SeuratWrappers::as.cell_data_set(data1)
cds <- cluster_cells(cds = cds, reduction_method = "UMAP",
                     resolution = 0.001) #saved 0.001
cds <- learn_graph(cds, 
                   use_partition = T, 
                   close_loop=T)

cds.Alldata = cds
save(cds.Alldata,file = 'Monocle.Alldata.newnames.RObj')
```
```{r, fig.height=8}
if(!(exists('cds.Alldata')))
load(file = 'Monocle.Alldata.newnames.RObj')
#* plot the results on the cluster graph. 
monocle3::plot_cells(cds.Alldata, 
                               color_cells_by = "ID.separate",
                               label_cell_groups=F, 
                               label_leaves=F,
                               label_branch_points=F,
                               label_roots = F,
                               trajectory_graph_color = 'black',
                               trajectory_graph_segment_size = 2,
                               cell_size = 1,
                               cell_stroke = 0,
)+ scale_color_manual(values = all.clusters.cp)

monocle3::plot_cells(cds.Alldata, 
             color_cells_by = "cytoTRACE.alldata",
             label_cell_groups=F, 
             label_leaves=F,
             label_branch_points=F,
             label_roots = F,
             trajectory_graph_color = 'cyan',
             trajectory_graph_segment_size = 1,
             cell_size = 1,
             cell_stroke = 0,
  )+viridis::scale_color_viridis(option = 'A',discrete = F,direction = -1,alpha = 0.5)
```
#Fig 8A
```{r}
pdf('Fig8A.pdf',width=8,height = 6)
monocle3::plot_cells(cds.Alldata, 
             color_cells_by = "cytoTRACE.alldata",
             label_cell_groups=F, 
             label_leaves=F,
             label_branch_points=F,
             label_roots = F,
             trajectory_graph_color = 'cyan',
             trajectory_graph_segment_size = 1,
             cell_size = 1,
             cell_stroke = 0,
  )+viridis::scale_color_viridis(option = 'A',discrete = F,direction = -1,alpha = 0.5)
```

## Celltype Tree Coarse
```{r}
library(stats)
library(ape)

# calculate a matrix of pseudobulk expression 
data1.ave <- Seurat::AverageExpression(Aa.Alldata,group.by = 'IDs',assays = 'RNA')
distance_matrix <- dist(t(data1.ave$RNA), method = "euclidean")
nj.tree <- nj(distance_matrix)

nj.boot.tree <- boot.phylo(nj.tree, t(data1.ave$RNA), FUN = function(xx) nj(dist(xx)), B = 1000)

# dividing by 10 to have a score between 0-100
nj.tree$node.label <- nj.boot.tree/10
# rounding because i don't want to plot decimal numbers
nj.tree$node.label <- round(nj.tree$node.label)
plot.phylo(nj.tree, type = "tidy", show.node.label = TRUE, label.offset = 0.1, rotate.tree = 90,tip.color = clust.cp.separate)
title(main="Cell tree | bootstrap replicates: 1000")


```
# image results: 
## UMAP library and cluster

```{r SubetPlots, eval=T}
clust.cp=clust.cp.separate
for (i in 1:length(names (data1.subsets)))
{
  data1=data1.subsets[[i]]
  print(
    DimPlot(data1,group.by = 'orig.ident',
            cols = LibCP) + NoAxes()+labs(title = 'Library | ID',subtitle = 'integrated')&
      DimPlot(data1, label = T,label.size = 5, repel = T,
              cols = clust.cp)+NoLegend()+NoAxes()+
      labs(title = names(data1.subsets)[i],subtitle = i)
  )
  
}
```

## Distribution barplots and cluster

```{r, eval=T,warning=FALSE,message=FALSE}

clust.cp=clust.cp.separate
for (i in 1:length(names (data1.subsets)))
{ 
  data1=data1.subsets[[i]]
  ids.cluster.library = as.data.frame(table(Idents(data1), data1@meta.data$orig.ident))
  colnames(ids.cluster.library) = c('ID','Library','CellCount')
  
  print(
    DimPlot(data1, label = F,label.size = 5, repel = T,
            cols = clust.cp)+NoLegend()+NoAxes()+
      labs(title = names(data1.subsets)[i],subtitle = i) +
      ggplot(ids.cluster.library, aes(fill=ID, y= log(CellCount),
                                      x=Library)) +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),
                            x=(Library)),
               position="stack", stat="identity", width = 0.5)+
      scale_fill_manual(values = clust.cp)+
      theme(axis.text.x = element_text(#face="bold", color="#993333", 
        size=8, angle=-45,hjust=0,vjust = 0.5))+
      geom_area(mapping =aes(fill=ID, y= log(CellCount),
                             x=as.integer(Library)),
                position="stack", stat="identity",alpha=0.2 , size=.5, colour="white") +
      geom_bar(mapping =aes(fill=ID, y= log(CellCount),#this re-plots the bars over the area
                            x=(Library)),
               position="stack", stat="identity", width = 0.5)+labs(title = names(data1.subsets)[i],subtitle = i)
  )
}
```


## DEG lists for subsets
- run once the refinement is finished and name are assigned...

```{r DEGs, warning=FALSE}
if(cluster.subsets)
{
  if (!exists('markers'))
    load(file = 'AllData.integrated.subsets.newnames.markers.RObj')
  if (!exists('markers.TF'))
    load(file = 'AllData.integrated.subsets.newnames.TF.markers.RObj')
  # markers = NULL
  # markers.TF = NULL# didn't work; off for now
  for (i in 1:length(names (data1.subsets)))
  {
    data1=data1.subsets[[i]]
    data1@active.assay='RNA'
    all.markers <- FindAllMarkers(data1,
                                  logfc.threshold = 0.6,
                                  # features = data1@misc$var.genes,
                                  return.thresh = 0.001,
                                  min.pct = 0.3,
                                  only.pos = TRUE, 
                                  max.cells.per.ident = 200,verbose=F)
    
    all.markers[,8:12] = 'NA'
    names(all.markers)[8:12]=names(Aa_genes)[c(1,3:6)]
    # add GO terms associated with this list:
    for (j in 1:length(levels(data1@active.ident))) # 
    {
      x=all.markers[as.numeric(all.markers$cluster)==j,][1:length(which(as.numeric(all.markers$cluster)==j)),7]
      x2=Aa_genes[match(x,Aa_genes$gene_short_name),c(1,3:6)]
      all.markers[as.numeric(all.markers$cluster)==j,][1:length(which(as.numeric(all.markers$cluster)==j)),8:12]<-x2
    }  
    markers[[i]] = all.markers 
    
    #also for transcription factors only: this did not work...
    all.markers_TF <- FindAllMarkers(data1,
                                     # logfc.threshold = 0.4,
                                     features = intersect(AaTF_list,rownames(data1)),
                                     min.pct = 0.01,
                                     only.pos = TRUE,
                                     max.cells.per.ident = 200,
                                     return.thresh = 0.001,verbose=F)
    
    all.markers_TF[,8:12] = 'NA'
    names(all.markers_TF)[8:12]=names(Aa_genes)[c(1,3:6)]
    # add GO terms associated with this list:
    for (j in 1:length(levels(data1@active.ident))) #
    {
      x=all.markers_TF[as.numeric(all.markers_TF$cluster)==j,][1:length(which(as.numeric(all.markers_TF$cluster)==j)),7]
      x2=Aa_genes[match(x,Aa_genes$gene_short_name),c(1,3:6)]
      all.markers_TF[as.numeric(all.markers_TF$cluster)==j,][1:length(which(as.numeric(all.markers_TF$cluster)==j)),8:12]<-x2
    }
    markers.TF[[i]] = all.markers_TF
  }
  # View(markers[[i]])
  names(markers) = names(data1.subsets)
  names(markers.TF) = names(data1.subsets)

  save(markers,file = 'AllData.integrated.subsets.newnames.markers.RObj')
  save(markers.TF,file = 'AllData.integrated.subsets.newnames.TF.markers.RObj')
}
```

plot the DEG lists
```{r}
 for (s in c(1:length(names (data1.subsets))))
  {
    list = NULL
    for (c in  1:length(levels(data1.subsets[[s]]@active.ident)))
    {
      x=markers[[s]][as.numeric(markers[[s]]$cluster)==c,][1:min(5,length(which(as.numeric(markers[[s]]$cluster)==c))),7]
      list=c(list,x)
    }
    
    print(DotPlot(data1.subsets[[s]],'RNA', features = unique(list),scale.by='radius' , col.min = 0, cols = c('lightgrey','darkred'))+ RotatedAxis() +FontSize(6,8)+ labs(title=names(data1.subsets)[s])+theme(axis.text.y = element_text(colour=all.clusters.cp[levels(data1.subsets[[s]])],size=8),legend.position = 'bottom'))
}
```
Plot the DETF lists
```{r}
 for (s in c(1:length(names (data1.subsets))))
  {
    list = NULL
    for (c in  1:length(levels(data1.subsets[[s]]@active.ident)))
    {
      x=markers.TF[[s]][as.numeric(markers.TF[[s]]$cluster)==c,][1:min(5,length(which(as.numeric(markers.TF[[s]]$cluster)==c))),7]
      list=c(list,x)
    }
    
    print(DotPlot(data1.subsets[[s]],'RNA', features = unique(list),
                              scale.by='radius' , col.min = 0, col.max = 4,  
                              cols = c('lightgrey','darkred'))+
      RotatedAxis() +FontSize(6,8)+ labs(title=names(data1.subsets)[s]))
}
```

### Cytotrace on subsets
```{r eval=F}
# load(file='AllData.integrated2.subsets.cyto.RObj') #saved here.
{
  for (i in c(1:length(names (data1.subsets)))) 
  {
    data1 = data1.subsets[[i]]
    print(FeaturePlot(data1,'cytoTRACE.subset', order=T, reduction = 'umap', cols = alpha(rev(brewer.pal(11 , "Spectral" )),0.5),label = T)+labs(title =names(data1.subsets[i]))+NoAxes()+NoLegend())
  }
}
```

#### save everything:
```{r SAVE.subsets, eval=F}
# save(data1.subsets,file = 'AllData.integrated2.subsets.RObj')
  save(data1.subsets,file = 'AllData.integrated.subsets.newnames.RObj')
```

# Add ID.separate to Alldata:
```{r}
if(cluster.subsets)
{
  cl.order=NULL
  cl.ind = NULL
  Aa.Alldata$IDs = Aa.Alldata@active.ident 
  Aa.Alldata$ID.separate = as.character(Aa.Alldata$IDs)
  for (i in c(1:5,8:10))#1:length(names (data1.subsets)))
  {
    coi=NULL
    coi=colnames(data1.subsets[[i]])
    Aa.Alldata$ID.separate[coi] = as.character(data1.subsets[[i]]@active.ident[coi])
    cl.order = c(cl.order,levels(data1.subsets[[i]]))
  }
  cl.ind = match(cl.order,levels(as.factor(Aa.Alldata$ID.separate))) 
  
  Aa.Alldata$ID.separate = as.factor(Aa.Alldata$ID.separate)
  Aa.Alldata$ID.separate = factor(Aa.Alldata$ID.separate,levels(Aa.Alldata$ID.separate)[cl.ind])
  
}
```

```{r, fig.height=12,eval=T}
print(DimPlot(Aa.Alldata,group.by = 'ID.separate',cols=all.clusters.cp)+NoAxes()+theme(legend.position = 'bottom'))
```
##Fig.2B

```{r, fig.height=12}

clust.cp=all.clusters.cp
data1=SetIdent(Aa.Alldata,value = 'ID.separate')
ids.cluster.library = as.data.frame(table(Idents(data1), data1@meta.data$lifehistory.tissue))
colnames(ids.cluster.library) = c('ID','Stage','CellCount')

print(
  ggplot(ids.cluster.library, aes(fill=ID, y= CellCount,
                                  x=Stage)) +
    geom_bar(mapping =aes(fill=ID, y= (CellCount),
                          x=(Stage)),
             position="fill", stat="identity", width = 0.5)+
    scale_fill_manual(values = clust.cp)+
    theme(axis.text.x = element_text(#face="bold", color="#993333", 
      size=8, angle=-45,hjust=0,vjust = 0.5))+
    geom_area(mapping =aes(fill=ID, y= (CellCount),
                           x=as.integer(Stage)),
              position="fill", stat="identity",alpha=0.2 , size=.5, colour="white") +
    geom_bar(mapping =aes(fill=ID, y= (CellCount),#this re-plots the bars over the area
                          x=(Stage)),
             position="fill", stat="identity", width = 0.5)+labs(title = 'All cluster identities')+theme(legend.position = 'none')
)
```
```{r}
pdf('Fig.2B.pdf',width = 6,height = 6)
ggplot(ids.cluster.library, aes(fill=ID, y= CellCount,
                                  x=Stage)) +
    geom_bar(mapping =aes(fill=ID, y= (CellCount),
                          x=(Stage)),
             position="fill", stat="identity", width = 0.5)+
    scale_fill_manual(values = clust.cp)+
    theme(axis.text.x = element_text(#face="bold", color="#993333", 
      size=8, angle=-45,hjust=0,vjust = 0.5))+
    geom_area(mapping =aes(fill=ID, y= (CellCount),
                           x=as.integer(Stage)),
              position="fill", stat="identity",alpha=0.2 , size=.5, colour="white") +
    geom_bar(mapping =aes(fill=ID, y= (CellCount),#this re-plots the bars over the area
                          x=(Stage)),
             position="fill", stat="identity", width = 0.5)+labs(title = 'All cluster identities')+theme(legend.position = 'none')
dev.off()
```
```{r}
save(Aa.Alldata,file='Aa.Alldata.newnames.RObj')
```

#### medusa-specific cell types:
```{r}
#generate indices for medusa specific cell types
data1=SetIdent(Aa.Alldata,value='ID.separate')

# maybe use the idents table:
ids.cluster.library = as.data.frame(table(Idents(data1), data1@meta.data$lifehistory))
colnames(ids.cluster.library) = c('ID','Lifehistory','CellCount')

ggplot(ids.cluster.library, aes(fill=ID, y= (CellCount-1),
                                x=Lifehistory)) +
    geom_bar(position="fill", stat="identity", width = 0.5)+
    scale_fill_manual(values = all.clusters.cp)+
      theme(axis.text.x = element_text(#face="bold", color="#993333", 
        size=8, angle=-45,hjust=0,vjust = 0.5))

cloi=ids.cluster.library$ID[which(ids.cluster.library$CellCount[ids.cluster.library$Lifehistory=='polyp'] <= 3)]
mcloi=ids.cluster.library$ID[which(ids.cluster.library$CellCount[!(ids.cluster.library$Lifehistory=='polyp' | ids.cluster.library$Lifehistory=='strobila')] <= 3)]

cl.oi=setdiff(levels(data1),cloi)
coi.medusa=WhichCells(data1,idents=cloi)
coi.all=WhichCells(data1,idents=cl.oi)
```

# image DEGs on full dataset
```{r, warning=F}
cluster.subsets = T
if(cluster.subsets)
{
  # Here write script to pull out the top 5 DEG's from the subsets and the cluster DEG's and plot onto the fully annotated clustered AllData
  if (!exists("Aa.Alldata"))
    load(file='AaAlldata.integrated.AAUR2.Robj')
  if (!exists("all.markers.Alldata"))
    load(file = 'AllData.integrated2.markers.RObj')
  if (!exists("data1.subsets"))
    load(file='AllData.integrated2.subsets.RObj')
  if (!exists("markers"))
    load(file='AllData.integrated2.subsets.markers.RObj')
  
  
  DEG.plot.all=NULL
  
  for (s in c(1:length(names (data1.subsets))))
    # s=2
  {
    list = NULL
    for (c in  1:length(levels(data1.subsets[[s]]@active.ident)))
    {
      x=markers[[s]][as.numeric(markers[[s]]$cluster)==c,][1:min(5,length(which(as.numeric(markers[[s]]$cluster)==c))),7]
      list=c(list,x)
    }
    
    DEG.plot.all[[s]]=DotPlot(Aa.Alldata,'RNA', #features = unique(list),
                              unique(c(list,all.markers.Alldata$gene[which(all.markers.Alldata$cluster == 
                                                                             names (data1.subsets)[s])][1:10])),
                              group.by = 'ID.separate',
                              scale.by='radius' , col.min = 0, col.max = 4,  
                              cols = c('lightgrey','darkred'))+
      RotatedAxis() +FontSize(6,8)+ labs(title=names(data1.subsets)[s])+theme(axis.text.y = element_text(colour=all.clusters.cp,size=8),legend.position = 'bottom')
  }
  for (s in c(1:length(names (data1.subsets))))
  {
    print(DEG.plot.all[[s]]) 
  }    
  # save(DEG.plot.all,file='DEG.onAll.Robj') #generates a crazy large file; in R is only 6.1MB
  
}
```
All of the DEGs ploted on the full dataset, with also top 10 of population markers.




```{r}
DETF.plot=NULL
 for (s in c(1:length(names (data1.subsets))))
  {
    list = NULL
    for (c in  1:length(levels(data1.subsets[[s]]@active.ident)))
    {
      x=markers.TF[[s]][as.numeric(markers.TF[[s]]$cluster)==c,][1:min(15,length(which(as.numeric(markers.TF[[s]]$cluster)==c))),7]
      list=c(list,x)
    }
    
    DETF.plot[[s]]=DotPlot(Aa.Alldata,'RNA', features = unique(list),
                              group.by = 'ID.separate',
                              scale.by='radius' , col.min = 0, col.max = 4,  
                              cols = c('lightgrey','darkred'),scale=T)+
      RotatedAxis() +FontSize(6,8)+ labs(title=names(data1.subsets)[s],subtitle = 'DETFs')+theme(axis.text.y = element_text(colour=all.clusters.cp,size=8))
  }
  for (s in c(1:length(names (data1.subsets))))
  {
    print(DETF.plot[[s]]) 
  }
```

These data suggest some mis-placements in the clustering (fuzzy boundaries and/or doublets)
*Questions that arise:*
1) Overlap with mature cnidocyte / mechanosensory. **Are there two separate things here?**
2) The cross cell type contamination: **doublets or cell decisions?**
3) Are n1.LWA and n1.2 neurons or digestive cells or doublets? 
- n1.2 genes are clearly in the dig. cluster and n1.LWA clearly has a 'polyp' signature
4) Mucin gland clearly has stMuscle contamination AND cnidocyte/muscle (mu.unchar.3)
- doublets should be looked at more carefully...
5) Inner cell layer has clear digestive gland overlaps; likely these originate here. **is this retained with the new approach?**
7) outer signal is strongest in cnidocytes / n2 neurons / mucin gland / striated muscle
8) n1.HLH5medusa has a st.muscle signature: **doublet?** This is of course feasible biologically becauase these are likely the medusa swimming moter neurons!

**Conclusions:** 
a) n1 + digestive gland clade derived from inner cell layer and shared across life history stages

#image cluster on full data
Not run...
```{r eval=T}
#* not run
#image results
clust.cp=clust.cp.separate
for (s in 1:length(names (data1.subsets)))
{
  data1=data1.subsets[[s]]
  cellID = NULL
  for (j in 1:length(levels(data1)))#unique(data2$seurat_clusters)))
    cellID[[j]] = WhichCells(data1, idents = levels(data1)[j])
  names(cellID)=levels(data1)
  
  print (DimPlot(Aa.Alldata, cells.highlight = cellID,
                 cols.highlight = alpha(clust.cp, 0.8),
                 label = F,pt.size = 1) +NoAxes())
  
}
```


```{r eval=F}
#  Plot the top 4 'marker' genes for each population on the subsets:
load(file = 'AllData.integrated2.markers.RObj')
data1=Aa.Alldata
for (i in 1:length(names (data1.subsets)))#i=19 fails at res 0.2 with fine clustering
{
  goi=all.markers.Alldata$gene[as.numeric(all.markers.Alldata$cluster)==i][1:4] 
  # data1.subsets[[i]]@active.assay = 'RNA'
  print(FeaturePlot(data1.subsets[[i]],goi,order=T,cols=gene.cp,label = T)&NoLegend()&NoAxes())
  # print(
  # DotPlot(Aa.Alldata,features=goi,idents = levels(Aa.Alldata)[i],split.by = 'orig.ident',cols=LibCP)
  # )
}

# save (data1.subsets,file = "subsets.single.1403.RObj")
```



```{r, fig.height=12}
data1=SetIdent(Aa.Alldata,value = 'ID.separate')
clust.cp=all.clusters.cp
ids.cluster.sample = as.data.frame(table(Idents(data1), data1@meta.data$lifehistory))
colnames(ids.cluster.sample) = c('ID', 'Lifehistory.Stage', 'CellCount')
ggplot(ids.cluster.sample, aes(fill = ID, y = CellCount,
                               x = Lifehistory.Stage)) +
  geom_bar(
    mapping = aes(
      fill = ID,
      y = (CellCount),
      x = (Lifehistory.Stage)
    ),
    position = "fill",
    stat = "identity",
    width = 0.5
  ) +
  scale_fill_manual(values = clust.cp) +
  theme(axis.text.x = element_text(
    #face="bold", color="#993333",
    size = 8,
    angle = -45,
    hjust = 0,
    vjust = 0.5
  )) +
  geom_area(
    mapping = aes(
      fill = ID,
      y = (CellCount),
      x = as.integer(Lifehistory.Stage)
    ),
    position = "fill",
    stat = "identity",
    alpha = 0.2 ,
    size = .5,
    colour = "white"
  ) +
  geom_bar(
    mapping = aes(
      fill = ID,
      y = (CellCount),
      #this re-plots the bars over the area
      x = (Lifehistory.Stage)
    ),
    position = "fill",
    stat = "identity",
    width = 0.5
  ) +
  ggtitle("Distribution of cell types over lifehistory stage")+
  theme(legend.position = 'NONE')
```

```{r}
data1=SetIdent(Aa.Alldata,value = 'ID.separate')
data1=subset(data1,idents=levels(data1)[c(1:3,24:length(levels(data1)))])
clust.cp=all.clusters.cp[c(1:3,24:length(levels(SetIdent(Aa.Alldata,value = 'ID.separate'))))]
ids.cluster.sample = as.data.frame(table(Idents(data1), data1@meta.data$lifehistory))
colnames(ids.cluster.sample) = c('ID', 'Lifehistory.Stage', 'CellCount')
ggplot(ids.cluster.sample, aes(fill = ID, y = CellCount,
                               x = Lifehistory.Stage)) +
  geom_bar(
    mapping = aes(
      fill = ID,
      y = (CellCount),
      x = (Lifehistory.Stage)
    ),
    position = "stack",
    stat = "identity",
    width = 0.5
  ) +
  scale_fill_manual(values = clust.cp) +
  theme(axis.text.x = element_text(
    #face="bold", color="#993333",
    size = 8,
    angle = -45,
    hjust = 0,
    vjust = 0.5
  )) +
  geom_area(
    mapping = aes(
      fill = ID,
      y = (CellCount),
      x = as.integer(Lifehistory.Stage)
    ),
    position = "stack",
    stat = "identity",
    alpha = 0.2 ,
    size = .5,
    colour = "white"
  ) +
  geom_bar(
    mapping = aes(
      fill = ID,
      y = (CellCount),
      #this re-plots the bars over the area
      x = (Lifehistory.Stage)
    ),
    position = "stack",
    stat = "identity",
    width = 0.5
  ) +
  ggtitle("Distribution of cell types over lifehistory stage")
``` 

# All IDs DEG plot from all.markers
```{r, warning=FALSE}
data1=SetIdent(Aa.Alldata,value = 'ID.separate')
   data1@active.assay='RNA'
    all.markers.allidents <- FindAllMarkers(data1,
                                  logfc.threshold = 1,
                                  # features = data1@misc$var.genes,
                                  return.thresh = 0.0001,
                                  min.pct = 0.5,
                                  only.pos = TRUE, 
                                  verbose=F)
    
    all.markers.allidents[,8:12] = 'NA'
    names(all.markers.allidents)[8:12]=names(Aa_genes)[c(1,3:6)]
    # add GO terms associated with this list:
    for (j in 1:length(levels(data1@active.ident))) # 
    {
      x=all.markers.allidents[as.numeric(all.markers.allidents$cluster)==j,][1:length(which(as.numeric(all.markers.allidents$cluster)==j)),7]
      x2=Aa_genes[match(x,Aa_genes$gene_short_name),c(1,3:6)]
      all.markers.allidents[as.numeric(all.markers.allidents$cluster)==j,][1:length(which(as.numeric(all.markers.allidents$cluster)==j)),8:12]<-x2
    }  
    
    #also for transcription factors only: this did not work...
    all.markers.TF.allidents <- FindAllMarkers(data1,
                                     logfc.threshold = 0.2,
                                     features = intersect(AaTF_list,rownames(data1)),
                                     min.pct = 0.05,
                                     only.pos = TRUE,
                                     max.cells.per.ident = 200,
                                     return.thresh = 0.001,verbose=F)
    
    all.markers.TF.allidents[,8:12] = 'NA'
    names(all.markers.TF.allidents)[8:12]=names(Aa_genes)[c(1,3:6)]
    # add GO terms associated with this list:
    for (j in 1:length(levels(data1@active.ident))) #
    {
      x=all.markers.TF.allidents[as.numeric(all.markers.TF.allidents$cluster)==j,][1:length(which(as.numeric(all.markers.TF.allidents$cluster)==j)),7]
      x2=Aa_genes[match(x,Aa_genes$gene_short_name),c(1,3:6)]
      all.markers.TF.allidents[as.numeric(all.markers.TF.allidents$cluster)==j,][1:length(which(as.numeric(all.markers.TF.allidents$cluster)==j)),8:12]<-x2
    }    
    
```

```{r, warning=FALSE}
list = NULL
    for (c in  1:length(levels(data1@active.ident)))
    {
      x=all.markers.allidents[as.numeric(all.markers.allidents$cluster)==c,][1:min(5,length(which(as.numeric(all.markers.allidents$cluster)==c))),7]
      list=c(list,x)
    }
        list2 = NULL
    for (c in  1:length(levels(data1@active.ident)))
    {
      x=all.markers.TF.allidents[as.numeric(all.markers.TF.allidents$cluster)==c,][1:min(5,length(which(as.numeric(all.markers.TF.allidents$cluster)==c))),7]
      list2=c(list2,x)
    }
         
    print(DotPlot(data1,'RNA', features = unique(list),
                              scale.by='radius' , col.min = 0, col.max = 4,  
                              cols = c('lightgrey','darkred'))+
      RotatedAxis() +FontSize(6,8) +labs(title='DEG')+theme(legend.position = 'bottom',legend.title = element_text(size=8),axis.text.y = element_text(colour=all.clusters.cp,size=8)))
      print(
        DotPlot(data1,'RNA', features = unique(list2),
                              scale.by='radius' , col.min = 0, col.max = 4,  
                              cols = c('lightgrey','darkred'))+
      RotatedAxis() +FontSize(6,8) +labs(title ='DETFs')+theme(axis.text.y = element_text(colour=all.clusters.cp,size=8)))
```

```{r}
save (all.markers.TF.allidents,all.markers.allidents,file='all.idents.markers.newnames.RData')
```


# Feature plots
```{r,fig.height=12,warning=FALSE,message=FALSE}
goi = c('twist','foxa2-like-1')
in.situ.genes=readxl::read_excel('Aurelia_genes_cloned.xlsx')
goi=genes$gene_short_name[match(in.situ.genes$`Gene ID`,genes$geneID)]
DotPlot(Aa.Alldata,'RNA',goi,group.by = 'ID.separate',cols=c('lightgrey','darkred'),col.min = 0)&RotatedAxis()&theme(axis.text.y = element_text(colour=all.clusters.cp,size=8),legend.position = 'bottom')
```
```{r}

```

```{r}
 data1$orig.ident = droplevels(as.factor(data1$orig.ident))
    ids.cluster.library = as.data.frame(table(Idents(data1), data1@meta.data$orig.ident))
    colnames(ids.cluster.library) = c('ID','Library','CellCount')


    print(ggplot(ids.cluster.library, aes(fill=ID, y= log(CellCount), x=Library)) + 
      geom_bar(position="stack", stat="identity")  +scale_fill_manual(values = clust.cp)+
      theme(axis.text.x = element_text(#face="bold", color="#993333", 
        size=8, angle=-45,hjust=0,vjust = 0.5)))
    
```
# General DEG 

```{r}
data1@active.assay='RNA'
    all.markers <- FindAllMarkers(data1,
                                  logfc.threshold = 1,
                                  # features = data1@misc$var.genes,
                                  return.thresh = 0.001,
                                  min.pct = 0.3,
                                  only.pos = TRUE, 
                                  verbose=F)
    
    all.markers[,8:12] = 'NA'
    names(all.markers)[8:12]=names(Aa_genes)[c(1,3:6)]
    # add GO terms associated with this list:
    for (j in 1:length(levels(data1@active.ident))) # 
    {
      x=all.markers[as.numeric(all.markers$cluster)==j,][1:length(which(as.numeric(all.markers$cluster)==j)),7]
      x2=Aa_genes[match(x,Aa_genes$gene_short_name),c(1,3:6)]
      all.markers[as.numeric(all.markers$cluster)==j,][1:length(which(as.numeric(all.markers$cluster)==j)),8:12]<-x2
    }  
      all.markers.TF <- FindAllMarkers(data1,
                                  logfc.threshold = 0.4,
                                  features = AaTF_list,
                                  return.thresh = 0.001,
                                  min.pct = 0.05,
                                  only.pos = TRUE, 
                                  verbose=F)
    
    all.markers.TF[,8:12] = 'NA'
    names(all.markers.TF)[8:12]=names(Aa_genes)[c(1,3:6)]
    # add GO terms associated with this list:
    for (j in 1:length(levels(data1@active.ident))) # 
    {
      x=all.markers.TF[as.numeric(all.markers.TF$cluster)==j,][1:length(which(as.numeric(all.markers.TF$cluster)==j)),7]
      x2=Aa_genes[match(x,Aa_genes$gene_short_name),c(1,3:6)]
      all.markers.TF[as.numeric(all.markers.TF$cluster)==j,][1:length(which(as.numeric(all.markers.TF$cluster)==j)),8:12]<-x2
    }  
    
```



```{r, warning=FALSE}
list = NULL
    for (c in  1:length(levels(data1@active.ident)))
    {
      x=all.markers[as.numeric(all.markers$cluster)==c,][1:min(5,length(which(as.numeric(all.markers$cluster)==c))),7]
      list=c(list,x)
    }
        list2 = NULL
    for (c in  1:length(levels(data1@active.ident)))
    {
      x=all.markers.TF[as.numeric(all.markers.TF$cluster)==c,][1:min(5,length(which(as.numeric(all.markers.TF$cluster)==c))),7]
      list2=c(list2,x)
    }
    
    print(DotPlot(data1,'alra', features = unique(list),
                              scale.by='radius' , col.min = 0, col.max = 4,  
                              cols = c('lightgrey','darkred'))+
      RotatedAxis() +FontSize(6,8) +labs(title='DEG')+theme(legend.position = 'bottom',legend.title = element_text(size=8),axis.text.y = element_text(colour=all.clusters.cp,size=8)))
      print(
        DotPlot(all.neuroglandular,'alra', features = unique(list2),
                              scale.by='radius' , col.min = 0, col.max = 4,  
                              cols = c('lightgrey','darkred'))+
      RotatedAxis() +FontSize(6,8) +labs(title ='DETFs')+theme(axis.text.y = element_text(colour=all.clusters.cp[levels(all.neuroglandular)],size=8)))
```

```{r,eval=FALSE,fig.height=12}
data1<-BuildClusterTree(SetIdent(Aa.Alldata,value='ID.separate'),
                        # assay = 'integrated',features = rownames(Aa.Alldata@assays$integrated))#,
                        # features = AaTF_list) # not useful
                        dims = 1:22)
# print(PlotClusterTree(data1,tip.color = all.clusters.cp,direction = 'rightwards'))
# data1@tools$BuildClusterTreeDims=data1@tools$BuildClusterTree
ape::plot.phylo(data1@tools$BuildClusterTree,type='tidy',no.margin = T,use.edge.length = F,tip.color = all.clusters.cp)
# ape::cophyloplot(data1@tools$BuildClusterTree,data1@tools$BuildClusterTreeDims)#didn't work
ape::plot.phylo(data1@tools$BuildClusterTree,type='fan',no.margin = T,use.edge.length = F,tip.color = all.clusters.cp)
ape::plot.phylo(data1@tools$BuildClusterTree,type='fan',no.margin = T,use.edge.length = T,tip.color = all.clusters.cp)

```

```{r,eval=FALSE,fig.height=12}
data1<-BuildClusterTree(SetIdent(Aa.Alldata,value='ID.separate'),
                        # assay = 'integrated',features = rownames(Aa.Alldata@assays$integrated))#,
                        features = AaTF_list) # not useful
                        # dims = 1:22)
# print(PlotClusterTree(data1,tip.color = all.clusters.cp,direction = 'rightwards'))
# data1@tools$BuildClusterTreeDims=data1@tools$BuildClusterTree
ape::plot.phylo(data1@tools$BuildClusterTree,type='tidy',no.margin = T,use.edge.length = F,tip.color = all.clusters.cp)
# ape::cophyloplot(data1@tools$BuildClusterTree,data1@tools$BuildClusterTreeDims)#didn't work
ape::plot.phylo(data1@tools$BuildClusterTree,type='fan',no.margin = T,use.edge.length = F,tip.color = all.clusters.cp)
ape::plot.phylo(data1@tools$BuildClusterTree,type='fan',no.margin = T,use.edge.length = T,tip.color = all.clusters.cp)

```

#Fig1E -----
```{r,eval=FALSE}
data1=SetIdent(Aa.Alldata,value='ID.separate')
    all.markers <- FindAllMarkers(data1,asssay='alra',
                                  logfc.threshold = 1.4,
                                  features = AaTF_list,
                                  return.thresh = 0.01,
                                  min.pct = 0.1,
                                  only.pos = TRUE, 
                                  max.cells.per.ident = 50,verbose=F)
    
    all.markers[,8:12] = 'NA'
    names(all.markers)[8:12]=names(Aa_genes)[c(1,3:6)]
    # add GO terms associated with this list:
    for (j in 1:length(levels(data1@active.ident))) # 
    {
      x=all.markers[as.numeric(all.markers$cluster)==j,][1:length(which(as.numeric(all.markers$cluster)==j)),7]
      x2=Aa_genes[match(x,Aa_genes$gene_short_name),c(1,3:6)]
      all.markers[as.numeric(all.markers$cluster)==j,][1:length(which(as.numeric(all.markers$cluster)==j)),8:12]<-x2
    }  
         all.markers.all.celltype=all.markers
save(all.markers.all.celltype,file='allcelltype.DETF')
```


```{r, fig.height=8,warning=FALSE}
Aa.Alldata=SetIdent(Aa.Alldata,value='ID.separate')
load(file='allcelltype.DETF')
#filter for p.value adjusted:
all.markers.all.celltype.filter = all.markers.all.celltype[all.markers.all.celltype$p_val_adj<=0.001,]
# neur.ids=c(47:64)
list = NULL
    for (c in (1:length(levels(Aa.Alldata@active.ident))))
    {
      x=all.markers.all.celltype.filter[as.numeric(all.markers.all.celltype.filter$cluster)==c,][1:min(3,length(which(as.numeric(all.markers.all.celltype.filter$cluster)==c))),7]
      list=c(list,x)
    }

    
    AllDETF.alldata=DotPlot(SetIdent(Aa.Alldata,value='ID.separate'),'RNA', features = unique(c(list,'ashc.2')),scale.by='size' , col.min = 0, cols = c('lightgrey','darkred'))+
      RotatedAxis() +FontSize(6,8) +labs(title='Top 3 DETFs / cell sub-type',subtitle = 'adj.p-val <= 0.001')+theme(legend.position = 'bottom',legend.title = element_text(size=8),legend.text = element_text(size=8),axis.text.y = element_text(colour=all.clusters.cp,size=9))
    
    print(AllDETF.alldata)
```

```{r}
pdf('Fig1E.pdf',width = 12,height = 6.5)
DotPlot(SetIdent(Aa.Alldata,value='ID.separate'),'RNA', features = unique(c(list)),scale.by='size' ,dot.min = 0.01 , col.min = 0, cols = c('grey90','darkred'))+
      RotatedAxis() +FontSize(6,6) +labs(title='Top 3 DETFs / cell sub-type',subtitle = 'adj.p-val <= 0.001')+theme(legend.position = 'bottom',legend.title = element_text(size=6),legend.text = element_text(size=6))
dev.off()
```
